### 2018-08-27 review/review.Rnw
Update the revision

\documentclass[a4paper]{article}
\usepackage[left=2.2cm, right=2.2cm,
            top=2.2cm, bottom=2.5cm]{geometry}

% Language
\usepackage[utf8]{inputenc}
\usepackage[english]{babel}

% Math packages
\usepackage{amsmath, amsthm, amsfonts, amssymb, amsxtra}
\usepackage{bm}

% Tables
\usepackage{booktabs}

% Fonts
%\usepackage{mathpazo}
\setlength\parindent{0pt}

% Utils
\usepackage{float}
\usepackage{lipsum}
\usepackage{hyperref}
\hypersetup{
  pdftitle={Reparametrization of the COM-Poisson Regression Models},
  colorlinks=true,
  linkcolor=red,
  citecolor=blue,
  filecolor=magenta,
  urlcolor=blue,
  bookmarksdepth=4
}

% Biliography
\usepackage{natbib}
\bibliographystyle{agsm}

% Metadata
\title{\vspace{-0.5cm}
  Reparametrization of COM-Poisson Regression Models with
  Applications in the Analysis of Experimental Data\\[0.5cm]
  \Large Replies to Reviewer's Comments
}
\author{
  Eduardo E. Ribeiro Jr\and
  Walmes M. Zeviani\and
  Wagner H. Bonat\and
  Clarice G. B. Dem\'{e}trio\and
  John Hinde}

%-----------------------------------------------------------------------
% Init document

<<include=FALSE>>=
knitr::read_chunk("review.R")
@

<<setup, echo=FALSE>>=
@

<<load-packages>>=
@

\begin{document}

\maketitle

\section{Replies to Referee 1}

\textbf{\underline{Overall reviewer's comment}:} \textit{Overall, the
  authors did a good job while presenting theirs results. However, I
  believe that the manuscript at the current state requires a
  considerable revision. Mainly, it requires a rigorous discussion with
  numerical results that when the proposed approach is useful and when
  it is not.}
\vspace{0.3cm}

\textbf{\underline{Response}:} We appreciate the referee for her/his
revision. All suggestions have been taken into account. Regarding the
referees' main comment, we propose a reparametrization of the
COM-Poisson model that is always valid. Only the interpretation of the
$\mu$ parameter depends on the approximation accuracy. We make it clear
in the updated version of the paper. The applicability of the model is
discussed at the end of Section 3 through the study of the dispersion
(DI), zero-inflation (ZI) and heavy-tail (HI) indexes (Figure 3).
\vspace{0.5cm}

\textbf{\underline{Specific reviewer's comments}:}
\begin{enumerate}
\item \textit{There has been some active research to find a better
    approximation for normalizing constant $Z(\lambda, \nu)$ from which
    the expressions for both mean and variance will be derived. Please
    refer to the Gaunt et al. (2016) for the latest results.}\\[0.1cm]
  \textbf{\underline{Response:}} Thanks for this reference. We have
  cited the paper published in the Annals of the Institute of
  Statistical Mathematics \cite{Gaunt2017}.
\item \textit{It seems that the paper missing some latest literature on
    CMP distribution. For example, the distributional properties of CMP
    were found in Daly and Gaunt (2015).}\\[0.1cm]
  \textbf{\underline{Response:}} Thanks for this reference. We have
  cited the paper published in the ALEA, Latin American Journal of
  Probability and Mathematical Statistics \cite{Daly2016}.
\item \textit{The likelihood equation (4.1) needs more details. Can we
    derive score equations? It seems like the likelihood has become even
    more complicated. Some of the nice properties of the CMP likelihood
    such as the canonical link are lost.}\\[0.1cm]
  \textbf{\underline{Response:}} The log-likelihood function cannot be
  derived in relation ti the dispersion parameter using any
  parametrization. We obtained the Hessian matrix by central finite
  differences using the Richardson method implemented in
  \texttt{R}. Although the expression of reparametrized log-likelihood
  (Equation 3.1) seems more complicated, this function is better behaved
  than the function expressed in the original form (see Figure
  8). Consider the exponential family of distributions, with probability
  function given by
  $$f(y) = \exp\{\sigma[\theta y - b(\theta)] - c(y, \phi)\},$$
  the COM-Poisson belongs to this family with $\sigma=1$,
  $\theta = \log(\lambda)$, $b(\theta) = \log(Z(\lambda, \nu))$ and
  $c(y, \phi)=-\nu\log(y!)$, only if $\nu$ is known. Since we don't have
  $b'(\theta) = \partial b(\theta)/\partial \theta$ in closed form, the
  canonical link function has no important rule here. In this case, the
  link function maps $\theta$ in the parameter space of $\lambda$.  In
  the reparametrized version of the COM-Poisson, the log link function
  maps the parameter space and provide simple interpretations (in terms
  of rate ratios).
\item \textit{For model fitting, the authors used general purpose
    optimization function in R. Did you observe any non convergence
    situations during your numerical studies.}\\[0.1cm]
  \textbf{\underline{Response:}} Yes, we have 37 situations where the
  algorithm did not converge in the simulation study (see table
  below). This occurred only for strong overdispersion and small sample
  sizes. We reported these situations in the updated version of the paper.
  \begin{table}[H]
    \centering
    \begin{tabular}{lcccc}
      \hline
      & $\phi=-1.6$ & $\phi=-1$ & $\phi=0$ & $\phi=1.8$ \\
      \hline
      $n=50$ &    29 &  0 &   0 &   0 \\
      $n=100$ &   8  &  0 &   0 &   0 \\
      $n=300$ &   0  &  0 &   0 &   0 \\
      $n=1000$ &  0  &  0 &   0 &   0 \\
      \hline
    \end{tabular}
  \end{table}
\item \textit{Numerical derivatives with CMP tend to be very slow and
    prone to errors. It would be good to have a computation times from
    the method for different sample sizes and dispersion
    levels.}\\[0.1cm]
  \textbf{\underline{Response:}} We have computed the times to fit the
  models for the case studies in the paper, they have different
  dispersion levels and sample sizes. However, the Figure below show the
  times to fit in the simulation study. The results show that it took
  more time to fit the models under equidispersion. This result is due
  to choose of start values and normalizing constant upper bound. We
  set, for all models, the upper bound 500 and we start the algorithm
  from the real parameters.

  \hspace*{-1cm}
\begin{minipage}{\textwidth}
<<simul-times, fig.width=9, fig.height=3.5>>=
@
\end{minipage}

\item \textit{In Section 5, the simulated counts are ranging from 3 to
    27. This is too limited. Please increase the range.}\\[0.1cm]
  \textbf{\underline{Response}:} The simulation study was performed with
  expectations varying from 3 to 27. Therefore, considering the
  different scenarios ($\phi=-1.6, \phi=-1, \phi=0$ and $\phi=1.8$), we
  cover a wide range of counts. The Figure below shows the variation of
  the counts for $\mu=27$ for the different scenarios.

  \hspace*{-1cm}
\begin{minipage}{\textwidth}
<<counts-ranging, fig.width=9, fig.height=4, cache=TRUE>>=
@
\end{minipage}

\item \textit{I believe that the authors simulated data according to the
    reparametrized distribution. Please generate the data according to
    the original distribution. Otherwise, the results not useful. For
    details on simulating from original CMP distribution, please refer
    to Chatla and Shmueli (2018).}\\[0.1cm]
  \textbf{\underline{Response:}} We simulated data according to the
  original parametrization. The steps in the simulation study were i)
  obtain the $\mu_i = \exp(\bm{x}_i^\top\bm{\beta})$; ii) back to the
  original parametrization $\lambda = (\mu + (\nu - 1)/(2\nu))^\nu$
  (Equation 3.1); and iii) simulate data according to the COM-Poisson
  distribution using the probability integral transform theorem as
  implemented in \texttt{compoisson} package \citep{compoisson2012}.

\item \textit{For real world data, the over dispersion case is too mild
    ($\phi=-0.77$). In practice over dispersion can be really high. For
    example, even lesser than $-1.4$. It would be nice to have such
    case, at least in the simulations.}\\[0.1cm]
  \textbf{\underline{Response:}} In the simulation study we consider the
  $\phi=-1.6$ and $\phi=-1$ that leads to dispersion indexes greater
  than four (see Figure 4).

\item \textit{From the presentation point of view, focusing more on
    applications would be good as there is not much novelty in the
    methodology.}\\[0.1cm]
  \textbf{\underline{Response:}} In this paper, we propose a
  mean-parametrization of the COM-Poisson model. This parametrization
  introduced $\mu = \lambda^{1/\nu} -(\nu-1)/(2\nu)$, a simple function
  of the original parameters $\lambda$ and $\nu$. We showed that the new
  parameter space has nice properties like i) the orthogonality between
  $\mu$ and $\phi$ (and consequently between $\beta_j$ and $\phi$ in a
  regression setting), leading to better computation and asymptotic
  (normal-based) inference \citep{Ross1970}; and ii) interpretability of
  the $\beta_j$ in terms of the rate ratios since
  $\mu_i\approx \text{E}(Y_i)$. Moreover, we characterized the
  COM-Poisson distribution in terms of the dispersion, zero-inflation
  and heavy-tail indexes that illustrate its applicability in the real
  count data. These are the methodological novelties of the paper
  besides the exploration of the results. However, in the updated
  version of the paper, we provided a better discussion of the case
  studies.
\end{enumerate}

\section{Replies to Referee 2}

\textbf{\underline{Second paragraph}:} \textit{The authors suggest the
  bijective reparameterization, $(\mu_i, \phi) = k(\lambda_i, \nu)$,
  where $\mu_i = \lambda_i^{1/\nu} - (\nu-1)/(2\nu)$. (Equation 3.1 plus
  the sentence after it gives this bijection. Note: The notation
  $h(\lambda, \nu)$, should probably be changed to $h_\nu(\lambda)$ to
  better hint that the symbol $h^{-1}$ refers to the inverse with
  respect to $\lambda$, with $\nu$ fixed.) The symbol $\mu_i$ is
  presumably used to remind the analyst that this parameter is, by
  Shmueli et al., approximately equal to $E(Y_i)$; we could write
  $m(\lambda_i, \nu) = E(Y_i) \approx \mu_i \equiv aE(Y_i)$. Shmueli et
  al. also showed that $var(Y_i)\approx \lambda_i^{1/\nu}\equiv a
  var(Y_i)$.}\\[0.1cm]
\textbf{\underline{Response:}} Thank you for this comment. We have
changed the notation of the reparameterization function in the updated
version.
\vspace{0.4cm}

\textbf{\underline{Third paragraph}:} \textit{The authors also hint that
  $var(Y_i)\approx \mu_i \exp(-\phi)$ (see the quadratic error analysis
  leading to Fig. 1). This is curious given that
  $avar(Y_i)=\mu\exp(-\phi) + (\exp(\phi)-1)/(2\exp(2\phi))$ and the
  latter summand converges to $-\infty$ as $\phi\to-\infty$. This latter
  summand is 0 or close to 0 for $\phi=0$ or $\phi>0$. Taken together,
  these last two sentences imply that the $\mu_i\exp(-\phi)$ may be a
  reasonable approximation to the variance when there is equi- or under-
  dispersion, but it may be unreasonable when there is lots of
  overdispersion. It seems that some of the results, simulations, and
  sample analyses reflect this (e.g. Fig 3, Fig 6-8, ``deviance function
  shape under strong overdispersion $\phi=-1.6$ is not as well
  behaved...'', p. 19, etc.). This needs to be addressed.}\\[0.3cm]
\textbf{\underline{Response:}} We thanks the referee for exploring this
point. However, we didn't use the approximation of variance
($a\text{Var}(Y_i)$) to reparametrize or fit the COM-Poisson model. To
compute the variances, that is showed in the Figure 3(a) and (b) we used
$\text{Var}(Y)=\sum_{y=0}^{500} y^2p(y) - [\sum_{y=0}^{500} y
p(y)]^2$.
As highlighted in the Figure 1(b), the approximation is not accurate,
mainly for overdispersion ($\phi<0$). The Figure (a) and (b) below shows
the behavior of $a\text{Var}(Y_i)$ function by fixing $\mu$ and $\phi$,
respectively, illustrating the referees' statement. The $a\text{Var}(Y)$
leads to negative values for variance for small $\mu$ and $\phi<0$ that
shows that the approximation is not reasonable for this region of the
parameter space.

\hspace*{-0.0cm}
\begin{minipage}{\textwidth}
<<approx-var, fig.width=9, fig.height=3.5, cache=TRUE>>=
@
\end{minipage}
\vspace{0.3cm}

The results shown in Figures 7 and 8 reflect that the orthogonality
property is slightly lost for strong overdispersion and small
$\mu$. This is related to the accuracy of the $aE(Y_i)$. When $aE(Y_i)$
is accurate, $\mu$ represents the expectation of the $Y_i$ and,
consequently, $\mu$ and $\phi$ (or $\beta_j$ and $\phi$ in a regression
setting) are orthogonal. We improved the discussion about this in the
current version of the paper.  \vspace{0.4cm}

\textbf{\underline{Fifth paragraph}:} \textit{This last point also hints
  that a quasi-Poisson model based on the assumption that
  $var(Y_i)\propto\mu_i$ (see p. 14) may not be reasonable for analyzing
  overdispersed data if the $COMPo$ model truly holds. Is this the
  case?}\\[0.1cm]
\textbf{\underline{Response:}} The quasi-Poisson model is specified by
second-moment assumptions (expectation and variance). As highlighted in
Figure 3(a) (and in referee's previous comment), the mean-variance
relationship for COM-Poisson is linear. Therefore, is suitable to
analyze data generated according to the COM-Poisson distribution, using
the assumption $\text{Var}(Y_i)\propto \mu_i$. The advantage of the
COM-Poisson approach is that it corresponds to a fully specified
probability model allowing to compute the likelihood and some useful
measures like deviance, AIC, BIC, LRT, etc.
\vspace{0.4cm}

\textbf{\underline{Sixth paragraph}:} \textit{Consider the example of
  Section 6.1. If I understand matters correctly, it is surprising that
  the $\beta$ estimates under the
  $[COMPo(\lambda_i, \nu), g(m(\lambda_i, \nu)) = \beta^\top x_i]$ model
  are so different from those under the
  $[COMPo(\mu_i, \phi), g(\mu_i) = \beta^\top x_i]$. See the results in
  Table 2, for example. I would think that if the approximation
  $m(\lambda_i, \nu) \approx \mu_i$ is reasonable, which your quadratic
  error results seem to indicate (see Fig 1), then the ML estimates of
  $\beta$ would be very similar. Exactly what models are being fitted
  here? This needs clarification.}\\[0.1cm]
\textbf{\underline{Response:}} Note that the $\lambda_i$ does not
represent the expectation, nor approximately. The quadratic errors in
Figure 1(a) is obtained by $[a\text{E(Y)} - \sum_{y=0}^{500}y p(y)]^2$.
The models fitted in the case studies are
$[\text{COMPo}(\lambda_i, \nu), \log(\lambda_i) = \beta^\top x_i]$ and
$[\text{COMPo}(\mu_i, \phi), \log(\mu_i) = \beta^\top x_i]$. Therefore,
the $\beta$'s estimates are comparable only if $\phi=0$ (Poisson special
case) otherwise they are on different scales. We make it clear in which
parameter the linear predictor is placed in the current version of the
paper. To understand the relationship of the coefficients, consider
$\log(\lambda_i) = \beta^o$ and $\log(\mu_i) = \beta^*$, so $\beta^*$
expressed in terms of $\beta^o$ is
$\beta^* = \exp(\beta^o-\phi) - (\exp(\phi) - 1)/(2\exp(\phi))$.
\vspace{0.4cm}

\textbf{\underline{Specific reviewer's comments}:}

\begin{enumerate}
\item \textit{Consider the constraint ``$\mu > 0$'', after equation
    (3.2). What constraints does this impose on the $\lambda$ and
    $\nu$. Is this reasonable?}\\[0.1cm]
  \textbf{\underline{Response:}} The constraint $\mu>0$ implies that
  $\lambda > [(\nu-1)/(2\nu)]^\nu$. The Figure below show (a) the
  constraint border in the parameter space; and (b) the expected values
  (obtained by $\sum y p(y)$). The constraint lies on the very small
  $\lambda$ and $\nu>1$ (underdispersion). This parameter region is
  related to small expected values (smaller than 0.1) and
  underdispersion. Therefore, although this constraint is undesirable,
  it does not prejudice the application of the model. We thank the
  referee for highlighting this point. We have discussed it in the
  updated version.

\hspace*{-1.0cm}
\begin{minipage}{\textwidth}
<<explore-constrain, fig.width=8, fig.height=4, cache=TRUE>>=
@
\end{minipage}
\vspace{0.3cm}

\item \textit{The Shmueli approximations to the mean and variance hold
    under certain conditions. Are those conditions met in practice? For
    your examples?}\\[0.1cm]
  \textbf{\underline{Response:}} We used only the mean approximation
  given by Shmueli to reparametrize the COM-Poisson model. Moreover, as
  discussed in Figure 1(a), the the errors are close to 0 for the
  parameter grid evaluated and present no clear relation with regions
  gives by \cite{Shmueli2005} ($\phi \leq 0$ and
  $\mu > 10 - (\exp(\phi) - 1)/(2\exp(\phi))$. Except for
  counts with small averages (close to 0), the reparametrized model fits
  well. For examples presented in the paper, the parameter estimates
  are summarizes below.

<<contraints-case-studies, results="asis", cache=TRUE>>=
@

\item \textit{The symbol $\mu$ is used to represent both $E(Y)$ and the
    parameter $\lambda^{1/\nu}-(\nu-1)/(2\nu) = aE(Y)$. This causes
    confusion. Is there a better alternative? As an example, in
    Simulation study of Section 5, at the bottom of p 14, does the
    symbol $\mu$ represent $E(Y)$ or $aE(Y)$? I assume the
    former.}\\[0.1cm]
  \textbf{\underline{Response:}} Thank you for advertising this
  misunderstanding. At the bottom of page 14 the $\bm{\mu}$ is
  $a\text{E}(\bm{Y})$. We keep symbol $\mu$ for the introduced parameter
  in order to highlight that it is related to the expectation. However,
  we make clear that $\mu_i = a\text{E}(Y_i)$ to avoid misunderstanding.

\item \textit{The abstract could be improved by making it more concise
    (shorter).  It should be written from the third-person perspective
    as well.}\\[0.1cm]
  \textbf{\underline{Response:}} We have written a more concise
  abstract. We keep it from a first-person perspective, like other
  papers published by the journal.

\item \textit{I assume that the “38\% faster” algorithm means that,
    e.g., instead of 10 seconds it takes just 6.2 seconds. As this is
    not an order of magnitude, this may not be enough to warrant a
    change to the “simpler” parameterization given that this latter
    parameterization models $aE(Y_i)$ rather than $E(Y_i)$.}\\[0.1cm]
  \textbf{\underline{Response:}} The time to fit the
  COM-Poisson$(\mu_i, \phi)$ models were compared to the time to fit the
  COM-Poisson$(\lambda_i, \phi)$ models in the paper. The computational
  times for 50 repetitions of fit of the models in the three case
  studies are presented in Figure below.

\hspace*{-1.0cm}
\begin{minipage}{\textwidth}
<<computation-times, fig.width=9, fig.height=3.5, cache=TRUE>>=
@
\end{minipage}
\vspace{0.3cm}

The computation times under the proposed parametrization are 110\%
faster than the original, in the overdispersed case. In addition, the
nice properties induced by the new parametrization like the
orthogonality between $\mu$ and $\phi$ and interpretation of $\mu$ are
the main advantages to warrant the use of the new parametrization
besides computational times.

\end{enumerate}

\bibliography{../references.bib}

\end{document}