### 2018-08-27 presentations/ibc2018/slides-ibc2018.Rnw

\documentclass[10pt]{beamer}

%=======================================================================
% General packages
\usepackage[british]{babel}
\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}
\usepackage{lmodern}
\usepackage{multicol}
\usepackage{mathtools} %% Funcionalidades (como \dcases)
\usepackage{dsfont}    %% Para \mathds{1} Indicadora
\usepackage{bm}

%=======================================================================
% For figures
\usepackage{graphicx}
\usepackage[export]{adjustbox}

%% ======================================================================
%% Fontes
\usepackage{mathpazo}
\usepackage{inconsolata}
\usepackage{verbatim}

\usefonttheme{professionalfonts}
\usefonttheme{serif}

%=======================================================================
% For tables
\usepackage[bottom]{footmisc}
\usepackage{booktabs}
\usepackage{threeparttable}
\usepackage{tabularx}
\usepackage{colortbl}
\newcolumntype{C}{>{\centering\arraybackslash}X}
\DeclareRobustCommand{\rchi}{{\mathpalette\irchi\relax}}
\newcommand{\irchi}[2]{\raisebox{\depth}{$#1\chi$}}
\usepackage{float}
\makeatletter
\def\@xfootnote[#1]{%
  \protected@xdef\@thefnmark{#1}%
  \@footnotemark\@footnotetext}
\makeatother

%% ======================================================================
%% Cores para links
\definecolor{url}{HTML}{000080}
\definecolor{run}{HTML}{4A0082}
\definecolor{col}{HTML}{6180A6}
\hypersetup{colorlinks, allcolors=., urlcolor=url, runcolor=run}

\setbeamercolor{bibliography entry author}{fg=black}
\usepackage{natbib}
\bibliographystyle{agsm}

%=======================================================================
% Algorithms
\usepackage[english, vlined, ruled]{algorithm2e}
\SetAlCapSty{}
\SetAlCapFnt{\footnotesize\color{col}}
\SetAlCapNameFnt{\footnotesize}

% Customize captions styles
\usepackage{caption}
\captionsetup{
  font=footnotesize,
  labelfont={color=col}
}

%% ======================================================================
%% Tema e cores do documento
\usetheme{CambridgeUS}
\setbeamertemplate{itemize items}[triangle]
\setbeamertemplate{navigation symbols}{}

\setbeamertemplate{frametitle}{
  \nointerlineskip
  \begin{beamercolorbox}[sep=0.3cm, ht=1.8em,
    wd=\paperwidth]{frametitle}
    \vbox{}\vskip-2ex%
    \strut\hspace*{3ex}\Large\bfseries\insertframetitle\strut
    \vskip-0.8ex%
  \end{beamercolorbox}
}

%% Slides em geral
\setbeamercolor{frametitle}{bg=white, fg=col}
\setbeamercolor{structure}{fg=col}
\setbeamercolor{palette primary}{bg=gray!30, fg=col}
\setbeamercolor{palette tertiary}{bg=col, fg=white}
\setbeamercolor{footlinecolor}{fg=white,bg=col}
\setbeamercolor{caption name}{fg=col}

% \setbeamertemplate{frametitle continuation}{[\insertcontinuationcount]}
\setbeamertemplate{frametitle continuation}{}

%% Slide Inicial
\setbeamertemplate{title page}[default]
\setbeamercolor{title}{fg=col}
\setbeamercolor{author}{fg=black!70}
\setbeamercolor{institute}{fg=black!70}
\setbeamercolor{date}{fg=black!70}
\setbeamerfont{title}{series=\bfseries, size=\Large}

%% ======================================================================
%% Definição do cabeçalho e rodapé
\setbeamerfont{headline}{size=\fontsize{5}{5}\selectfont}
\setbeamertemplate{headline}{\bfseries
  \leavevmode%
  \hbox{%
    \begin{beamercolorbox}[wd=.5\paperwidth, ht=2.2ex, dp=1ex, right,
      rightskip=1em]{section in head/foot}
      \hspace*{2ex}\insertsectionhead
    \end{beamercolorbox}%
    \begin{beamercolorbox}[wd=.5\paperwidth, ht=2.2ex, dp=1ex, left,
      leftskip=1em]{subsection in head/foot}
      \insertsubsectionhead\hspace*{2ex}
    \end{beamercolorbox}}
  \vskip0pt
}

\setbeamerfont{footline}{size=\fontsize{5}{5}\selectfont}
\makeatletter
\setbeamertemplate{footline}{\bfseries
  \leavevmode%
  \hbox{%
    \begin{beamercolorbox}[wd=.3\paperwidth, ht=2.4ex, dp=1ex, right,
      rightskip=1em]{footlinecolor}
      \centering \insertshortauthor%
    \end{beamercolorbox}%
    \begin{beamercolorbox}[wd=.6\paperwidth, ht=2.4ex, dp=1ex, left,
      leftskip=1em]{footlinecolor}
      \hfill\insertshorttitle%
    \end{beamercolorbox}%
    \begin{beamercolorbox}[wd=.1\paperwidth, ht=2.4ex, dp=1ex, left,
      leftskip=1em]{footlinecolor}
      \centering Slide \insertframenumber
    \end{beamercolorbox}}
  \vskip0pt
}
\makeatother

%% ======================================================================
%% Layout do tableofcontents
\setbeamertemplate{section in toc}{
  {\color{col} \bfseries\inserttocsectionnumber.}~
  {\leftskip=0.5em\color{black}\inserttocsection\par}
}

\setbeamertemplate{subsection in toc}{
  {\color{col!80}
  \bfseries\inserttocsectionnumber.\inserttocsubsectionnumber}~
  \leftskip=2em{\color{black}\inserttocsubsection\par}
}

%% ======================================================================
%% Formatando slides para seções e subseções
\AtBeginSection[]{
  \begin{frame}[c, allowframebreaks, noframenumbering]
    \begin{center}
      \textcolor{col}{\thesection} \\ \vspace{0.3cm}
      \parbox{0.6\textwidth}{
        \centering \textcolor{col}{\LARGE \bf \insertsection}}\\
    \end{center}
  \end{frame}
}

\AtBeginSubsection{
  \begin{frame}[c, allowframebreaks, noframenumbering]
    \begin{center}
      \textcolor{col}{\thesection.\thesubsection} \\ \vspace{0.3cm}
      \parbox{0.6\textwidth}{
        \centering \textcolor{col!80}{\large \insertsection}\\
        \centering \textcolor{col}{\Large \bf \insertsubsection}}\\
    \end{center}
  \end{frame}
}

%% ======================================================================
%% Metadados não disponíveis no YAML com o atual default.tex template

\title[Reparametrization of COM-Poisson Regression Models]{
  Reparametrization of COM-Poisson Regression Models with Applications
  in the Analysis of Experimental Count Data \vspace{-0.4cm}}

\author[Clarice Garcia Borges Demétrio]{
  Eduardo Elias Ribeiro Junior \inst{1} \inst{2} \\
  Walmes Marques Zeviani \inst{1} \\
  Wagner Hugo Bonat \inst{1} \\
  Clarice Garcia Borges Demétrio \inst{2} \\
  John Hinde \inst{3}
}

\institute{
  \inst{1} Statistics and Geoinformation Laboratory (LEG-UFPR) \\
  \inst{2} Department of Exact Sciences (ESALQ-USP) \\
  \inst{3} School of Mathematics, Statistics and Applied Mathematics (NUI-Galway)
}

\date{\small 10th July 2018 \\
  \url{jreduardo@usp.br} | \url{edujrrib@gmail.com}}

<<setup, include=FALSE>>=

library(knitr)
library(xtable)
options(digits = 3, OutDec = ".",
        xtable.caption.placement = "top",
        xtable.booktabs = TRUE,
        xtable.sanitize.text.function = identity,
        xtable.sanitize.colnames.function = identity)

opts_chunk$set(
    warning = FALSE,
    message = FALSE,
    echo = FALSE,
    results = "asis",
    ## dev = "tikz",
    fig.width = 7,
    fig.height = 5,
    out.width = "0.99\\textwidth",
    fig.align = "center",
    fig.pos = "h",
    dev.args = list(family = "Palatino"))

thm = knit_theme$get("dusk")
knit_theme$set(thm)

@

\begin{document}

\begin{frame}[plain, noframenumbering]
  \begin{center}
    \bfseries
    XXIX\textsuperscript{TH} INTERNATIONAL BIOMETRIC CONFERENCE\\
    Barcelona, Spain, 8-13 July 2018
  \end{center}
  \vspace{-0.5cm}
  \titlepage
\end{frame}

\begin{frame}[noframenumbering]{Outline}
  \tableofcontents[hideallsubsections]
\end{frame}

\section{Background}

\begin{frame}{Count data}

  Number of times an event occurs in the observation unit.\\[0.3cm]

  Random variables that assume non-negative integer values.\\[0.3cm]

  Let $Y$ be a counting random variable, so that $y=0,1,2,\ldots$\\[0.5cm]

  Examples in experimental researches:

  \begin{itemize}
    \item number of grains produced by a plant;
    \item number of fruits produced by a tree;
    \item number of insects on a particular cell;
    \item others.
  \end{itemize}

\end{frame}

\begin{frame}{Poisson model and limitations}

  \begin{itemize}
    \item[] {\bf GLM framework \citep{Nelder1972}}
    \item Provide suitable distribution for a counting random variables;
    \item Efficient algorithm for estimation and inference;
    \item Implemented in many software.
      \vspace{0.5cm}
    \item[] {\bf Poisson model}
    \item Relationship between mean and variance,
      $\text{E}(Y)=\text{Var}(Y)$;
      \vspace{0.5cm}
    \item[] {\bf Main limitations}
    \item Overdispersion (more common),
      $\text{E}(Y)<\text{Var}(Y)$
    \item Underdispersion (less common),
      $\text{E}(Y)>\text{Var}(Y)$
  \end{itemize}

\end{frame}

\begin{frame}{COM-Poisson distribution}

  \begin{itemize}
    \itemsep7.5pt\parskip0pt\parsep0pt
  \item Probability mass function \citep{Shmueli2005} takes the form \\
    \begin{equation*} \small
    \label{eqn:pmf-cmp}
    \Pr(Y = y \mid \lambda, \nu) =
      \frac{\lambda^y}{(y!)^\nu Z(\lambda, \nu)}\,,
    \qquad Z(\lambda, \nu) = \sum_{j=0}^\infty
      \frac{\lambda^j}{(j!)^\nu}\,,
  \end{equation*}
    where $\lambda > 0$ and $\nu \geq 0$.
  \item Moments are not available in closed form;
  \item Expectation and variance can be closely approximated by
    \begin{equation*}
      \label{eqn:mean-aprox}
      \text{E}(Y) \approx \lambda^{1/\nu} - \frac{\nu - 1}{2\nu} \qquad
      \textrm{and} \qquad
      \text{Var}(Y) \approx \frac{\lambda^{1/\nu}}{\nu}
    \end{equation*}
    with accurate approximations for $\nu \leq 1$ or $\lambda > 10^\nu$
    \citep{Shmueli2005, Sellers2012}.
  \end{itemize}

\end{frame}

\begin{frame}{COM-Poisson regression models}

\begin{itemize}
  \item[] {\bf Model definition}
  \item Modelling the relationship between
    $\text{E}(Y_i)$ and $\bm{x}_i$ indirectly
    \citep{Sellers2010}; \\[-0.5cm]
    \begin{align*}
      Y_i \mid & \,\bm{x}_i \sim
                 \textrm{COM-Poisson}(\lambda_i,\, \nu) \\
               &\eta(E(Y_i \mid \bm{x}_i)) =
                 \log(\lambda_i) = \bm{x}_i^\top\bm{\beta} &
    \end{align*}
\end{itemize}
\vspace{0.3cm}

\begin{itemize}
  \item[] {\bf Main goals}
  \item Study distribution properties in terms of i) modelling real
    count data and ii) inference aspects.
  \item Propose a reparametrization in order to model the expectation of
    the response variable as a function of the covariate values
    directly.
\end{itemize}

\end{frame}

\section{Reparametrization}

\begin{frame}{Reparametrized COM-Poisson}

\begin{itemize}
  \item[] {\bf Reparametrization}
  \item Introduced new parameter $\mu$, using the mean
    approximation\\[-0.2cm]
    \begin{equation*}
      \label{eqn:repar-cmp}
      \mu = \lambda^{1/\nu} - \frac{\nu - 1}{2\nu}
      \quad \Rightarrow \quad
      \lambda = \left (\mu +
        \frac{(\nu - 1)}{2\nu} \right )^\nu;
    \end{equation*}
  \item Precision parameter is taken on the log scale to avoid
    restrictions on the parameter space\\[-0.3cm]
    $$\phi = \log(\nu) \Rightarrow \phi \in \mathbb{R};$$\\[0.3cm]
  \item[] {\bf Probability mass function}
  \item Replacing $\lambda$ and $\nu$ as function of $\mu$ and $\phi$ in
    the pmf of COM-Poisson\\[-0.2cm]
    \begin{equation*}
      \label{eqn:pmf-cmpmu}
      \Pr(Y=y \mid \mu, \phi) =
      \left ( \mu +\frac{ e^\phi-1}{2e^\phi} \right )^{ye^\phi}
      \frac{(y!)^{-e^\phi}}{Z(\mu, \phi)}.
    \end{equation*}
\end{itemize}

\end{frame}

\begin{frame}{Study of the moments approximations}

<<approx-plot, fig.height=3, fig.width=6, out.width=".9\\textwidth", fig.cap="Quadratic errors for the approximation of the (a) expectation and (b) variance. Dotted lines represent the restriction for suitable approximations given by \\cite{Shmueli2005}.">>=
include_graphics("figure/approx-plot-1.pdf")
@

\end{frame}

\begin{frame}{COM-Poisson$_\mu$ distribution}

<<pmf-cmp, fig.height=3, fig.width=7, fig.cap="Shapes of the COM-Poisson distribution for different parameter values.">>=
include_graphics("figure/pmf-cmp-1.pdf")
@

\end{frame}

\begin{frame}{Properties of COM-Poisson distribution}
  To explore the flexibility of the COM-Poisson distribution, we
  consider the follow indexes:
  \begin{itemize}
  \item \textbf{Dispersion index:}\quad\:\;
    $\text{DI} = \text{Var}(Y)/\text{E}(Y);$\\
  \item \textbf{Zero-inflation index:}\;
    $\text{ZI} = 1 + \log \Pr(Y = 0)/\text{E}(Y);$
  \item \textbf{Heavy-tail index:}\quad\:\;\,
    $\text{HT} = \Pr(Y=y+1)/\Pr(Y=y),\:
    \text{for}\: y \to \infty.$\\
  \end{itemize}
  \vspace{0.4cm}

  These indexes are interpreted in relation to the Poisson distribution:
  \begin{itemize}
  \item
    over- ($\text{DI} > 1$), under- ($\text{DI} < 1$) and
    equidispersion ($\text{DI} = 1$);\\
  \item
    zero-inflation ($\text{ZI} > 0$) and zero-deflation ($\text{ZI} <
    0$) and
  \item
    heavy-tail distribution for $\text{HT} \to 1$ when $y \to
    \infty$.
  \end{itemize}
\end{frame}

\begin{frame}{Properties of COM-Poisson distribution}

<<indexes-plot, fig.width=9.5, fig.height=4, fig.cap="Indexes for COM-Poisson distribution. (a) Mean and variance relationship, (b--d) dispersion, zero-inflation and heavy-tail indexes for different parameter values. Dotted lines represents the Poisson special case.">>=
include_graphics("figure/indexes-plot-1")
@

\end{frame}

\begin{frame}{COM-Poisson$_\mu$ regression models}

\begin{itemize}
\item[] Let $y_i$ a set of independent observations from the COM-Poisson
  and $\bm{x}_i^\top = (x_{i1},\, x_{i2},\, \ldots,\, x_{ip})$ is a
  vector of known covariates, $i=1,2,\ldots,n$.
  \vspace{0.3cm}

\item[] {\bf Model definition}
\item Modelling relationship between $E(Y_i)$ and $\bm{x}_i$
  directly\\[-0.5cm]
  \begin{align*}
    Y_i \mid & \,\bm{x}_i \sim
               \textrm{COM-Poisson}_\mu(\mu_i,\, \phi) \\
             &\log(E(Y_i \mid \bm{x}_i)) =
               \log(\mu_i) = \bm{x}_i^\top\bm{\beta} &
  \end{align*}
  \vspace{-0.2cm}
\item[] {\bf Log-likelihood function
    ($\ell = \ell(\bm{\beta},\, \phi \mid \bm{y})$)}
  \vspace{0.05cm}
\item
  $
    \begin{aligned}[t]
    &\ell = e^\phi \left [
      \sum_{i=1}^n y_i
      \log \left( \mu_i + \frac{e^\phi-1}{2e^\phi} \right )
      -\sum_{i=1}^n \log(y_i!) \right ]
      -\sum_{i=1}^n \log(Z(\mu_i, \phi))\\
      &\text{\small where }\:
      \mu_i = \exp(\bm{x}_i^\top\bm{\beta})
      \end{aligned}
  $
\end{itemize}

\end{frame}

\begin{frame}{Estimation and inference}

  The estimation and inference is based on the method of maximum
  likelihood. Let $\bm{\theta} = (\bm{\beta}, \phi)$ the model
  parameters. \vspace{0.2cm}
  \begin{itemize}
  \item Parameter estimates are obtained by numerical maximization of
    the log-likelihood function (by BFGS algorithm);\\[0.1cm]
    $\ell(\hat{\bm{\theta}}) =
    \text{max}\,\ell(\bm{\theta}),\:
    \bm{\theta} \in \mathbb{R}^{p+1}$;
    \vspace{0.2cm}
  \item Standard errors for regression coefficients are obtained based
    on the observed information matrix;\\[0.1cm]
    $\text{Var}(\hat{\bm{\theta}}) = -\mathcal{H}^{-1}$, where
    $\mathcal{H}$ is the matrix of second partial derivatives at
    $\hat{\bm{\theta}}$; \vspace{0.2cm}
  \item Confidence intervals for $\hat{\mu}_i$ are obtained by delta
    method.\\[0.1cm]
    $\text{Var}[g(\hat{\bm{\theta}})] \doteq \bm{G}\,
    \text{Var}(\hat{\bm{\theta}}) \bm{G}^\top $,
    where
    $\bm{G}^\top = (\partial g / \partial \beta_1,
    \ldots, \partial g / \partial \beta_p)^\top$;
  \item The Hessian matrix $\mathcal{H}$ is obtained numerically by
    finite differences.
  \end{itemize}

\end{frame}

\section{Simulation study}

\begin{frame}{Definitions on the simulation study}
  \small

  \textbf{Objective:} assess the properties of maximum likelihood
  estimators and orthogonality in the reparametrized model;
  \vspace{0.2cm}

  \textbf{Simulation:} we consider counts generated according a
  regression model with a continuous and categorical covariates and
  different dispersion scenarios.
  \vspace{-0.1cm}

  \begin{center}
  \begin{minipage}{0.9\textwidth}
  \begin{algorithm}[H]
    \footnotesize
    \caption{Steps in simulation study.}
      \For{$n \in \{50, 100, 300, 1000\}$}{ set $\bm{x}_1$ as a
        sequence, with $n$ elements, between $0$ and $1$\; set
        $\bm{x}_2$ as a repetition, with $n$ elements, of three
        categories\;
        compute $\bm{\mu}$ using
        $\bm{\mu} = \exp(\beta_0 + \beta_1 \bm{x}_1 + \beta_{21}
        \bm{x}_{21} + \beta_{22} \bm{x}_{22})$\;
        \For{$\phi \in \{-1.6, -1.0, 0.0, 1.8\}$}{
          \Repeat{$1000$ times}{
            simulate $\bm{y}$ from COM-Poisson distribution with
            $\bm{\mu}$ and $\phi$ parameters\;
            fit COM-Poisson$_\mu$ regression model to simulated
            $\bm{y}$\;
            get $\hat{\bm{\theta}} = (\hat{\phi},\;
              \hat{\beta}_0,\; \hat{\beta}_1,\; \hat{\beta}_{21},\;
              \hat{\beta}_{22})$\;
            get confidence intervals for $\hat{\bm{\theta}}$ based on
            the observed information matrix.
          }
        }
      }
  \end{algorithm}
  \end{minipage}
  \end{center}
\end{frame}

\begin{frame}{Definitions on the simulation study}

<<justpars, fig.height=4, fig.width=9, fig.cap="Average counts (left) and dispersion indexes (right) for each scenario considered in the simulation study.">>=
include_graphics("figure/justpars-1")
@

\end{frame}

\begin{frame}{Bias of the estimators}

\vspace{-0.5cm}
<<bias-plot, out.width="0.9\\textwidth", fig.height=6.5, fig.width=8.5, fig.cap="Distributions of standardized bias (gray box-plots) and average with confidence intervals (black segments) by different sample sizes and dispersion levels.">>=
include_graphics("figure/bias-plot-1")
@

\end{frame}

\begin{frame}{Coverage rate of the confidence intervals}

<<coverage-plot, fig.height=4, fig.width=9, fig.cap="Coverage rate based on confidence intervals obtained by quadratic approximation for different sample sizes and dispersion levels.">>=
include_graphics("figure/coverage-plot-1")
@

\end{frame}

\begin{frame}{Orthogonality property of the MLEs}
\vspace{-0.3cm}
<<ortho-surf, fig.height=5, fig.width=9, fig.cap="Deviance surfaces contour plots under original and proposed parametrization. The ellipses are confidence regions (90, 95 and 99\\%), dotted lines are the maximum likelihood estimates, and points are the real parameters used in the simulation.">>=
include_graphics("figure/ortho-surf-1")
@

\end{frame}

\section{Case studies}

\begin{frame}[t]{Motivating data sets and data analysis}

  \begin{itemize}
    \item Three illustrative examples of count data analysis are
      reported.
      \begin{itemize} \normalsize
      \item Assessing toxicity of nitrofen in aquatic systems, an
        equidispersed example;
      \item Soil moisture and potassium doses on soybean culture, an
        overdispersed example; and
      \item Artificial defoliation in cotton phenology, an
        underdispersed example.
      \end{itemize}
    \item In the data analysis, we consider the COM-Poisson model in the
      two forms (original and new parametrization) and the quasi-Poisson
      regression model as alternative models for the standard Poisson
      regression model.
  \end{itemize}
\end{frame}

\subsection{Artificial defoliation in cotton phenology}

\begin{frame}{Cotton bolls data}

  \begin{columns}
    \begin{column}{0.4\textwidth}
      \begin{figure}
        \includegraphics[scale=0.25]{./images/defoliation}
      \end{figure}
    \end{column}
    \begin{column}{0.6\textwidth}
      {\bf Aim:} to assess the effects of five defoliation levels on the
      bolls produced at five growth stages;\\[0.2cm]
      {\bf Design:} factorial 5 $\times$ 5, with 5 replicates;\\[0.2cm]
      {\bf Experimental unit:} a plot with 2 plants;\\[0.2cm]
      {\bf Factors:}
      \begin{itemize} \small
        \item Artificial defoliation (\texttt{des}):
        \item Growth stage (\texttt{est}):
      \end{itemize}\vspace{0.1cm}
      {\bf Response variable:} Total number of cotton bolls;\\
    \end{column}
  \end{columns}

\end{frame}

\begin{frame}{Model specification}

  {\bf Linear predictor:} {\small following \citet{Zeviani2014}}
  \begin{itemize}
    \item $\log(\mu_{ij}) = \beta_0 + \beta_{1j} \textrm{def}_i +
      \beta_{2j} \textrm{def}_i^2$\\[0.1cm]
    {\small
    $i$ varies in the levels of artificial defoliation;\\
    $j$ varies in the levels of growth stages.}
  \end{itemize}

  \vspace{0.5cm}

  {\bf Alternative models:}
  % {\bf Modelling strategies:}
  \begin{itemize}
    \item Poisson ($\mu_{ij}$);
    \item COM-Poisson ($\lambda_{ij}=\eta(\mu_{ij})\,,\,\phi$)
    \item COM-Poisson$_\mu$ ($\mu_{ij}\,,\,\phi$)
    \item Quasi-Poisson (var$(Y_{ij})=\sigma\mu_{ij}$)
  \end{itemize}

\end{frame}

\begin{frame}{Parameter estimates}

  \vspace{-0.6cm}
\begin{table}[!ht]
\centering \footnotesize
\caption{Parameter estimates (Est) and ratio between estimate and
  standard error (SE).}
\vspace{-0.1cm}
\label{tab:coef-cotton}
\begin{tabular}{lrrrrrrrr}
  \toprule
  & \multicolumn{2}{c}{Poisson} &
    \multicolumn{2}{c}{COM-Poisson} &
    \multicolumn{2}{c}{COM-Poisson$_\mu$} &
    \multicolumn{2}{c}{Quasi-Poisson} \\
\cmidrule(lr){2-3} \cmidrule(lr){4-5} \cmidrule(lr){6-7} \cmidrule(lr){8-9}
% latex table generated in R 3.4.4 by xtable 1.8-2 package
% Tue Jun 26 13:34:54 2018
 & Est & Est/SE & Est & Est/SE & Est & Est/SE & Est & Est/SE \\
  \midrule
  \rowcolor{col!30}
$\phi\,,\,\sigma$ &  &  & 1.585 & 12.417 & 1.582 & 12.392 & 0.241 &  \\
  $\beta_0$ & 2.190 & 34.572 & 10.897 & 7.759 & 2.190 & 74.640 & 2.190 & 70.420 \\
  $\beta_{11}$ & 0.437 & 0.847 & 2.019 & 1.770 & 0.435 & 1.819 & 0.437 & 1.726 \\
  $\beta_{12}$ & 0.290 & 0.571 & 1.343 & 1.211 & 0.288 & 1.223 & 0.290 & 1.162 \\
  $\beta_{13}$ & $-$1.242 & $-$2.058 & $-$5.750 & $-$3.886 & $-$1.247 & $-$4.420 & $-$1.242 & $-$4.192 \\
  $\beta_{14}$ & 0.365 & 0.645 & 1.595 & 1.298 & 0.350 & 1.328 & 0.365 & 1.314 \\
  $\beta_{15}$ & 0.009 & 0.018 & 0.038 & 0.035 & 0.008 & 0.032 & 0.009 & 0.036 \\
  $\beta_{21}$ & $-$0.805 & $-$1.379 & $-$3.725 & $-$2.775 & $-$0.803 & $-$2.961 & $-$0.805 & $-$2.809 \\
  $\beta_{22}$ & $-$0.488 & $-$0.861 & $-$2.265 & $-$1.805 & $-$0.486 & $-$1.850 & $-$0.488 & $-$1.754 \\
  $\beta_{23}$ & 0.673 & 0.989 & 3.135 & 2.084 & 0.679 & 2.135 & 0.673 & 2.015 \\
  $\beta_{24}$ & $-$1.310 & $-$1.948 & $-$5.894 & $-$3.657 & $-$1.288 & $-$4.095 & $-$1.310 & $-$3.967 \\
  $\beta_{25}$ & $-$0.020 & $-$0.036 & $-$0.090 & $-$0.076 & $-$0.019 & $-$0.074 & $-$0.020 & $-$0.074 \\
   \specialrule{0.01em}{0.3em}{0.3em}
 LogLik & \multicolumn{2}{c}{$-255.803$} & \multicolumn{2}{c}{$-208.250$} & \multicolumn{2}{c}{$-208.398$} & \multicolumn{2}{c}{$  -$}\\
 AIC & \multicolumn{2}{c}{$533.606$} & \multicolumn{2}{c}{$440.500$} & \multicolumn{2}{c}{$440.795$} & \multicolumn{2}{c}{$  -$}\\
 BIC & \multicolumn{2}{c}{$564.718$} & \multicolumn{2}{c}{$474.440$} & \multicolumn{2}{c}{$474.735$} & \multicolumn{2}{c}{$  -$} \\
 \bottomrule
\end{tabular}
\end{table}
\end{frame}

\begin{frame}{Fitted curves}

<<pred-cotton, fig.height=3.5, fig.width=8.5, fig.cap="Scatterplots of the observed data and curves of fitted values with 95\\% confidence intervals as functions of the defoliation level for each growth stage.">>=
include_graphics("figure/pred-cotton-1")
@

\end{frame}

\subsection*{Additional results}

\begin{frame}{Computational times and orthogonality property}

  To compare the computational times on the two parametrizations we
  repeat the fitting 50 times.
  \vspace{-0.2cm}

<<comp-times, fig.height=4, fig.width=8, fig.cap="Computational times to fit the models under original and reparametrized versions based on the fifty repetitions.">>=
include_graphics("figure/comp-times-1")
@

\end{frame}

\section{Final remarks}

\begin{frame}{Concluding remarks}

  {\bf Summary}

  \begin{itemize}
    \item Over/under-dispersion needs caution;
    \item COM-Poisson is a suitable choice for these situations;
    \item The proposed reparametrization, COM-Poisson$_\mu$ has some
      advantages:
      \begin{itemize}
        \item Simple transformation of the parameter space;
        \item Leads to the orthogonality of the parameters (seen
          empirically);
        \item Full parametric approach;
        \item Empirical correlation between the estimators was
          practically null;
        \item Faster for fitting;
        \item Allows interpretation of the coefficients directly (like
          GLM-Poisson model).
      \end{itemize}
  \end{itemize}

  {\bf Future work}

  \begin{itemize}
    \item Simulation study to assess model robustness against
      distribution miss specification;
    \item Assess theoretical approximations for $Z(\lambda, \nu)$
      (or $Z(\mu, \phi)$), in order to avoid the selection of sum's
      upper bound;
    \item Propose a mixed GLM based on the COM-Poisson$_\mu$ model.
  \end{itemize}
\end{frame}

\section*{Notes}

\begin{frame}

\begin{itemize}
  \item \begin{columns}[c]
      \column{.1\linewidth}
      \begin{flushright}
        \includegraphics[height=0.9cm,valign=c]{./images/arxiv}
      \end{flushright}
      \column{.8\linewidth}
      \hspace*{-0.7cm} Full-text article is available on arXiv\\
      \hspace*{-0.7cm} \url{https://arxiv.org/abs/1801.09795}
    \end{columns}
  \item \begin{columns}[c]
      \column{.1\linewidth}
      \begin{flushright}
        \includegraphics[height=0.9cm,valign=c]{./images/github-logo}
      \end{flushright}
      \column{.8\linewidth}
      \hspace*{-0.7cm} All codes (in R) and source files are available
      on GitHub\\
      \hspace*{-0.7cm} \url{https://github.com/jreduardo/article-reparcmp}
    \end{columns}
\end{itemize}

\vspace{0.5cm}

{\bf Acknowledgements}

\begin{itemize}
    \item \begin{columns}[c]
      \column{.1\linewidth}
      \begin{flushright}
        \includegraphics[height=0.9cm,valign=c]{./images/cnpq-logo}
      \end{flushright}
      \column{.8\linewidth} \vspace{-0.3cm}
      \hspace*{-0.7cm}\begin{minipage}[c]{\textwidth}
       National Council for Scientific and Technological Development
       (CNPq), for their support.\\
      \end{minipage}
    \end{columns}
\end{itemize}

\end{frame}

\section*{Bibliography}
\begin{frame}[allowframebreaks]{References}
  \small
  \vspace*{-0.5cm}
  \bibliography{../../references.bib}
\end{frame}

\end{document}