### 2018-09-13 article.Rnw
Make the changes suggested by John Hinde

%-------------------------------------------
% Document class
\documentclass[publish]{smj}

%-------------------------------------------
% Personal packages and definitions
\usepackage{amsfonts}
\usepackage{float}
\usepackage[section]{placeins}
\usepackage{bm}
\DeclareRobustCommand{\rchi}{{\mathpalette\irchi\relax}}
\newcommand{\irchi}[2]{\raisebox{\depth}{$#1\chi$}}

% Algorithm
\usepackage[english, vlined, ruled]{algorithm2e}
\SetAlCapSty{}

\usepackage{booktabs}
\usepackage{tabularx}
\newcolumntype{C}{>{\centering\arraybackslash}X}

% -------------------------------------------
% Authors and affiliations
\Author{Eduardo E. Ribeiro Jr\Affil{1},
  Walmes M. Zeviani\Affil{2},
  Wagner H. Bonat\Affil{2},
  Clarice G. B. Dem\'{e}trio\Affil{1}
  and John Hinde\Affil{3}
}
\AuthorRunning{Eduardo E. Ribeiro Jr \textrm{et al.}}

\Affiliations{
\item Department of Exact Sciences,
  University of S\~{a}o Paulo - ESALQ,
  Piracicaba,
  SP, Brazil
\item Department of Statistics,
  Paran\'{a} Federal University,
  Curitiba,
  PR, Brazil
\item School of Mathematics, Statistics and Applied Mathematics,
  National University of Ireland Galway,
  Galway, Ireland
}

%-------------------------------------------
% Corresponding author
\CorrAddress{Eduardo E. Ribeiro Jr,
             Department of Exact Sciences,
             Luiz de Queiroz College of Agriculture - ESALQ,
             University of S\~{a}o Paulo - USP,
             Piracicaba, S\~{a}o Paulo,
             P\'{a}dua Dias, 11, Avenue, CEP--13.418-900
             Brazil}
\CorrEmail{jreduardo@usp.br}
\CorrPhone{(+55)\;41\;9\;8711\;9034}
\CorrFax{(+55)\;19\;3447\;6021}

%-------------------------------------------
% Title and short title
\Title{Reparametrization of COM-Poisson Regression Models with
  Applications in the Analysis of Experimental Data}
\TitleRunning{Reparametrization of COM-Poisson Regression Models}

% -------------------------------------------
% Abstract
\Abstract{The COM-Poisson distribution is a two-parameter generalization
  of the Poisson distribution that can deal with under-, equi- and
  overdispersed count data. Unfortunately, its location parameter does
  not correspond to the expectation, which complicates the
  interpretation of regression models specified using this
  distribution. In this paper, we propose a straightforward
  reparametrization of the COM-Poisson distribution based on an
  approximation to the expectation. The main advantage of our new
  parametrization is the interpretation of the regression coefficients
  in terms of the (approximate) expectation of the count response
  variable, as is usual in the context of generalized linear
  models. Estimation and inference for the reparametrized COM-Poisson
  regression models can be done using the likelihood paradigm. We
  carried out simulation studies to verify the finite sample properties
  of the maximum likelihood estimators. The results from our simulation
  study show that the maximum likelihood estimators are unbiased and
  consistent for both regression and dispersion parameters. The
  empirical correlation between the regression and dispersion parameter
  estimators and nature of the deviance surfaces suggests that these
  parameters are also approximately orthogonal, which is advantageous
  for interpretation, inference, and computational efficiency. We
  illustrate the application of the proposed model through the analysis
  of three data sets with over-, under- and equidispersed count
  data. The study of the distribution's properties, through a
  consideration of dispersion, zero-inflation and heavy tail indexes,
  together with the results of data analysis, show the flexibility over
  standard approaches.  The computational routines for fitting the
  original and reparameterized versions of the COM-Poisson regression
  model and data sets are available in the supplementary material.  }

%-------------------------------------------
% Key-words
\Keywords{ COM-Poisson, Count data, Likelihood inference,
  Overdispersion, Underdispersion. }

<<include=FALSE>>=
knitr::read_chunk("article.R")
@

<<setup, include=FALSE>>=
@

%-----------------------------------------------------------------------

\begin{document}

\maketitle

\section{Introduction}
\label{introduction}

Count data arise from random variables that take non-negative integer
values and typically represent the number of times an event occurs in an
observation period or region. This kind of data is also common in crop
sciences, examples including the number of grains produced by a plant,
the number of fruits on a tree, the number of insects captured by a
trap, etc.  Since the seminal paper of \citet{Nelder1972}, where the
class of the generalized linear models (GLMs) was introduced, the
analysis of count data often uses a Poisson regression model. This model
provides a suitable strategy for the analysis of count data and an
efficient Newton scoring algorithm that can be used for fitting.

In spite of the advantages of the Poisson regression model, the Poisson
distribution has only one parameter that represents both the expectation
and variance of the count random variable.  This equality restriction on
the relationship between the expectation and variance of the Poisson
distribution is referred as equidispersion.  However, in practical data
analysis such restriction can be unsuitable, since the observed data can
present variance both smaller or larger than the mean, leading to under-
and overdispersion, respectively.  While applying the Poisson regression
model to non-equidispersed count data can give consistent parameter
estimates, the associated standard errors are inconsistently estimated,
which in turn can lead to misleading inferences~\citep{Winkelmann1995,
  Bonat2017}.

In practice, overdispersion is widely reported in the literature and may
occur due to the absence of relevant covariates, heterogeneity of
sampling units, different observational periods/regions not being
considered in the analysis, and excess zeros~\citep{Hinde1998}.  The
case of underdispersion is less often reported in the literature,
however, it has been of increasing interest in the statistical
community. The processes that reduce the variability are not as
well-known as those leading to extra variability.  For this reason,
there are few models to deal with underdispersed count data. Possible
explanatory mechanisms leading to underdispersion may be related to the
underlying stochastic process generating the count data.  For example,
when the time between events is not exponentially distributed, the
number of events can be over or underdispersed; a process that motivated
the class of duration dependence models~\citep{Winkelmann1995}.  Another
possible explanation for underdispersion is when the responses
correspond to order statistics of component observations, such as maxima
of Poisson distributed counts~\citep{Steutel1989}.

Strategies for constructing alternative count distributions are related
to the causes of the non-equidispersion. Specifically for
overdispersion, Poisson mixture (compound) models are widely applied.
One popular example of this approach is the negative-binomial model,
where the expectation of the Poisson distribution is assumed to be gamma
distributed.  However, other distributions can also be used to represent
this additional random variation.  For example the Poisson-Tweedie
model~\citep{Bonat2017} and its special cases the Poisson
inverse-Gaussian and Neyman-Type A models assume that the random effects
are Tweedie, inverse Gaussian or Poisson distributed, respectively.  The
Gamma-Count distribution assumes a gamma distribution for the time
between events and it can deal with underdispersed as well as
overdispersed count data~\citep{Zeviani2014}. The subject of this paper,
the COM-Poisson distribution, is obtained by a generalization of the
Poisson distribution that allows for a non-linear decrease in the ratios
of successive probabilities~\citep{Shmueli2005}.

The COM-Poisson distribution includes the Poisson and geometric
distributions as special cases, as well as the Bernoulli distribution as
a limiting case. It can deal with both under- and overdispersed count
data.  Some recent applications of the COM-Poisson distribution include
\citet{Lord2010} for the analysis of traffic crash data,
\citet{Sellers2010} for the modelling of airfreight breakage and book
purchases, \citet{Huang2017} on the analysis of attendance data,
takeover bids and cotton boll counts, and \citet{Chatla2018} to model
counts from bike sharing systems. Theoretical results and approximations
derived for this distribution are discussed by \citet{Shmueli2005},
\citet{Daly2016} and \citet{Gaunt2017}. The main disadvantage of the
COM-Poisson regression model as presented in~\citet{Sellers2010} is that
its location parameter does not correspond to the expectation of the
distribution. This complicates the interpretation of regression models
and means that they are not comparable with standard approaches, such as
the Poisson and negative binomial regression models. In order to avoid
this issue, \citet{Huang2017} proposed a mean-parametrization of the
COM-Poisson distribution.  In his approach the mean parameter is
obtained as the solution of a non-linear equation defined as an infinite
sum, which is computationally demanding and liable to numerical
problems.

The main goal of this article is to propose a novel COM-Poisson
parametrization based on the mean approximation presented by
\citet{Shmueli2005}. In this parametrization, the probability mass
function is written in terms of $\mu$ and $\phi$, where $\mu$ is the
approximate expectation and $\phi$ is a dispersion parameter. In
contrast to the original parametrization, the proposed parametrization
leads to regression coefficients directly associated (approximately)
with the expectation of the response variable, as is usual in the
context of generalized linear models.  Consequently, the resulting
regression coefficients are comparable to those obtained by standard
approaches, such as the Poisson and negative binomial regression models.
Furthermore, our novel COM-Poisson parametrization is computationally
simpler than the strategy proposed by~\citet{Huang2017}, since it does
not require any numerical method for solving non-linear equations. Also,
we show that attractive properties like the orthogonality between
dispersion and regression parameters and consistency and asymptotic
normality of the maximum likelihood estimators are retained.

This paper is organized as follows. In \autoref{background} we present
the COM-Poisson distribution and the approach proposed
by~\citet{Huang2017}.  The newly proposed reparametrization is
considered in ~\autoref{reparametrization}, along with assessment of the
moment approximation, and a study of distribution's
properties. In~\autoref{estimation-and-inference} we present estimation
and inference for the novel COM-Poisson regression model in a likelihood
framework. The properties of the maximum likelihood estimators and their
approximate orthogonality are assessed in~\autoref{simulation-study}
through simulation studies.  We illustrate the application of the new
COM-Poisson regression model with the analysis of three data sets. We
provide an \texttt{R} implementation for the COM-Poisson and
reparameterized COM-Poisson regression models, together with the
analyzed data sets, in the supplementary material.\footnote{Available on
  \texttt{\url{http://www.leg.ufpr.br/~eduardojr/papercompanions}}
  \label{papercompanion}.}.

\section{Background}
\label{background}

The COM-Poisson distribution generalizes the Poisson distribution in
terms of the ratio between the probabilities of two consecutive events
by adding an extra dispersion parameter~\citep{Sellers2010}.  Let $Y$ be
a COM-Poisson random variable, then
$$\frac{\Pr(Y=y-1)}{\Pr(Y=y)} = \frac{y^\nu}{\lambda}$$ while for the
Poisson distribution this ratio is $\frac{y}{\lambda}$ corresponding to
$\nu=1$.  This allows the COM-Poisson distribution to deal with
non-equidispersed count data.  The probability mass function of the
COM-Poisson distribution is given by
\begin{equation}
  \label{eqn:pmf-cmp}
  \Pr(Y=y \mid \lambda, \nu) = \frac{\lambda^y}{(y!)^\nu Z(\lambda, \nu)},
  \qquad y = 0, 1, 2, \ldots,
\end{equation}
where $\lambda > 0$, $\nu \geq 0$ and
$Z(\lambda, \nu) = \sum_{j=0}^\infty \frac{\lambda^j}{(j!)^\nu}$ is a
normalizing constant that depends on both parameters.

The $Z(\lambda, \nu)$ series diverges theoretically only when $\nu=0$
and $\lambda \geq 1$, but numerically, for small values of $\nu$
combined with large values of $\lambda$, the sum is so huge it results
in overflow. \autoref{tab:convergenceZ} shows the values of the
normalizing constant based on one thousand terms in the summation, that
is $\sum_{j=0}^{1000}\lambda^j/(j!)^\nu$, for different values of
$\lambda$ and $\phi$.

<<convergenceZ, echo=FALSE>>=
@

\begin{table}[ht]
\centering
\caption{Values for $Z(\lambda, \nu)$ normalizing constant (computed numerically with 1000 terms in the summation) for values of $\lambda$ (0.5 to 50) and $\phi$ (0 to 1)}
\label{tab:convergenceZ}
\begingroup\small
\begin{tabularx}{\textwidth}{C|CCCCCC}
  \toprule

<<results-Z, results="asis">>=
@

\end{tabularx}
\endgroup

\footnotesize \raggedright
divergent$^{*}$ denotes a mathematically divergent series; and
divergent$^{**}$ a numerically divergent series.
\end{table}

In the first line of \autoref{tab:convergenceZ} we have mathematically
divergent series, because $\sum_{j=0}^\infty\lambda^j$ is divergent when
$\lambda \geq 1$.  In other cases the series diverges numerically, due
to the computational storage limitation.  For both forms of divergence
it is impossible to compute COM-Poisson probabilities, therefore, this
places a restriction on the feasible parameter space.

An undesirable feature of the COM-Poisson distribution is that the
moments cannot be obtained in closed form. \citet{Shmueli2005} and
\citet{Sellers2010} using an asymptotic approximation for
$Z(\lambda,\nu)$, showed that the expectation and variance of the
COM-Poisson distribution can be approximated by
\begin{equation}
  \label{eqn:mean-approx}
  \text{E}(Y) \approx \lambda^{1/\nu} - \frac{\nu - 1}{2\nu} \qquad
  \textrm{and} \qquad
  \textrm{Var}(Y) \approx \frac{\lambda^{1/\nu}}{\nu}\,,
\end{equation}with greatest accuracy for  $\nu \leq 1$ or
$\lambda > 10$. The authors also argue that the mean-variance
relationship can be approximated as
$\textrm{Var}(Y) \approx\frac{1}{\nu}\text{E}(Y)$.
In~\autoref{reparametrization}, we assess the accuracy of these
approximations.

The COM-Poisson regression model was proposed by~\citet{Sellers2010},
using this original parametrization. Specifically, the COM-Poisson
regression model is defined as
$\log(\lambda_i) = \bm{x}_i^\top \bm{\beta}$ and the relationship
between E$(Y_i)$ and $\bm{x}_i$ is modelled indirectly.
\citet{Huang2017} shows how to model directly the expectation of the
COM-Poisson distribution in a suitable reparametrization. From
\autoref{eqn:pmf-cmp}, \Citeauthor{Huang2017} notes that the parameter
$\lambda$ is given, as a function of $\mu$ and $\nu$, by the solution to
\begin{equation*}
  \sum_{j=0}^{\infty} (j - \mu) \frac{\lambda^j}{(j!)^\nu} = 0\,.
\end{equation*}
This allows \Citeauthor{Huang2017} to define a mean-parametrized
COM-Poisson regression model with
$\log(\mu_i) = \bm{x}_i^\top \bm{\beta}$ giving a direct relationship
between $\mu$ and the covariates $\bm{x}$. In this article, we propose
an alternative approximate mean-parametrization of the COM-Poisson
distribution in order to avoid the limitations of the original
parametrization and the numerical complexity of the
\Citeauthor{Huang2017}'s approach.

\section{Reparametrized COM-Poisson regression model}
\label{reparametrization}

The proposed reparametrization of COM-Poisson models is based on the
mean approximation (\autoref{eqn:mean-approx}). We introduce a new
parameter $\mu$, using this approximation,
\begin{equation}
  \label{eqn:repar-cmp}
  \mu = h_\nu(\lambda) = \lambda^{1/\nu} - \frac{\nu - 1}{2\nu}
  \quad \Rightarrow \quad
  \lambda = h_\nu^{-1}(\mu) = \left (\mu +
    \frac{(\nu - 1)}{2\nu} \right )^\nu.
\end{equation}
The dispersion parameter is taken on the log scale for computational
convenience, thus $\phi = \log(\nu)$, $\phi \in \mathbb{R}$. The
interpretation of $\phi$ is the same as the $\nu$, but simply on another
scale. For $\phi < 0$ and $\phi > 0$ we have the overdispersed and
underdispersed cases, respectively.  When $\phi=0$ we have the Poisson
distribution as a special case.

<<data-approx, include=FALSE, cache=TRUE>>=
@

In order to assess the accuracy of the moment approximations
(\autoref{eqn:mean-approx}), \autoref{fig:approx-plot} presents the
squared errors for (a) expectation and (b) variance.  The squared errors
are $[\mu - \text{E}(Y)]^2$ for the expectation and
$[ \mu \exp(-\phi) - \text{Var}(Y)]^2$ for the variance. In both cases
$\text{E}(Y)$ and Var$(Y)$ were computed numerically using 500
terms\footnote{ E$(Y) \approx \sum_{y=0}^{500}yp(y)$ and
  E$(Y^2) \approx \sum_{y=0}^{500}y^2p(y)$.}. The dotted lines represent
the border between the regions $\nu \leq 1$ and $\lambda > 10^\nu$, in
the $\mu$ and $\phi$ scale.

<<approx-plot, fig.height=3, fig.width=6, out.width=".8\\textwidth", fig.cap="Squared errors for the approximation of the (a) expectation and (b) variance. Dotted lines represent the restriction for suitable approximations given by \\cite{Shmueli2005}.">>=
@

The results in \autoref{fig:approx-plot} show that the mean
approximation is accurate, the largest squared error, for the parameter
values evaluated, is $0.038$.  For the variance approximation, the
largest squared error was $XXX$ and it occurs for negative values of
$\phi$. Interestingly, the errors for the variance are larger for
negative values of $\phi$ and present no clear relation with $\mu$, as
opposed to the regions gives by \citet{Shmueli2005} ($\phi \leq 0$ and
$\mu > 10 - \frac{\exp(\phi) - 1}{2\exp(\phi)}$).

The results presented in \autoref{fig:approx-plot}(a) support the
proposed reparametrization. Replacing $\lambda$ and $\nu$ as functions of
$\mu$ and $\phi$ in \autoref{eqn:pmf-cmp}, the reparametrized
distribution takes the form
\begin{equation}
  \label{eqn:pmf-cmpmu}
  \Pr(Y=y \mid \mu, \phi) =
  \left ( \mu +\frac{ e^\phi-1}{2e^\phi} \right )^{ye^\phi}
  \frac{(y!)^{-e^\phi}}{Z(\mu, \phi)},
  \qquad y = 0, 1, 2, \ldots\,,
\end{equation}
where $\mu > 0$. We denote this reparameterized distribution as
COM-Poisson$_\mu$. In \autoref{fig:pmf-cmp}, we show the shapes of
the COM-Poisson$_\mu$ distribution.

<<pmf-cmp, fig.height=3, fig.width=7, fig.cap="Shapes of the COM-Poisson distribution for different parameter values.">>=
@

The constraint $\mu>0$ imposes an undesirable restriction in the
original parameter space, $\lambda^\nu>(\nu-1)/(2\nu)$. However, this
restricted region is related to small underdispersed counts (averages
smaller than 0.1 and $\nu>1$) a parameter region unlikely to be of
interest in practice.

In order to explore the flexibility of the COM-Poisson model to deal
with real count data, we compute indexes for dispersion (DI),
zero-inflation (ZI) and heavy-tail (HI), which are respectively given by
\begin{equation*}
\text{DI} = \frac{\text{Var}(Y)}{\text{E}(Y)}, \quad
\text{ZI} = 1 + \frac{\log \Pr(Y = 0)}{\text{E}(Y)}
  \quad \text{and} \quad
\text{HT} = \frac{\Pr(Y=y+1)}{\Pr(Y=y)}\quad \text{for} \quad y \to
\infty.
\end{equation*}
These indexes are defined in relation to the Poisson distribution. Thus,
the dispersion index indicates overdispersion for $\text{DI} > 1$,
underdispersion for $\text{DI} < 1$ and equidispersion for
$\text{DI} = 1$. The zero-inflation index indicates zero-inflation for
$\text{ZI} > 0$, zero-deflation for $\text{ZI} < 0$ and no excess of
zeros for $\text{ZI} = 0$. Finally, the heavy-tail index indicates a
heavy-tail distribution for $\text{HT} \to 1$ when $y \to \infty$.
These indexes are discussed by \citet{Bonat2017} to study the
flexibility of Poisson-Tweedie distribution, and \citet{Puig2006} to
describe count distributions in general.

<<compute-indexes, include=FALSE, cache=TRUE>>=
@

<<indexes-plot, fig.width=9.5, fig.height=4, fig.cap="Indexes for the COM-Poisson distribution. (a) Mean and variance relationship, (b--d) dispersion, zero-inflation and heavy-tail indexes for different parameter values. Dotted lines represents the special case of the Poisson distribution.">>=
@

Regarding the COM-Poisson$_\mu$ distribution, in
\autoref{fig:indexes-plot} we present the relationship between (a) mean
and variance, (b--c) the dispersion and zero-inflation indexes for
different values of $\mu$ and $\phi$, and (d) the heavy-tail index for
$\mu=25$ and different values of $y$ and $\phi$.
\autoref{fig:indexes-plot} shows that the indexes are slightly dependent
on the expected values and tend to stabilize for large values of
$\mu$. Consequently, the mean and variance relationship
\autoref{fig:indexes-plot}(a) is proportional to the dispersion
parameter $\phi$. In terms of moments, this leads to a specification
indistinguishable from the quasi-Poisson regression model. The
dispersion indexes in \autoref{fig:indexes-plot}(b) show that the
distribution is suitable to deal to dispersed counts, of course. For the
parameter values evaluated the largest DI was $XXX$ and smallest was
$XXX$. \autoref{fig:indexes-plot}(c) shows the COM-Poisson can handle
a limited amount of zero-inflation, in cases of overdispersion
($\phi < 0$). On the other hand, for $\phi > 0$ (underdispersion) this
distribution is suitable to deal with zero-deflated counts. Heavy-tail
indexes in \autoref{fig:indexes-plot}(d) indicate the distribution is in
general a light-tailed distribution, i.e. $\text{HT} \to 0$ for
$y \to \infty$.

\section{Estimation and Inference}
\label{estimation-and-inference}

In this section we describe estimation and inference for the two forms
of the COM-Poisson regression model based on the maximum likelihood
method. Inference can be done using the standard machinery of likelihood
inference, including likelihood ratio tests for model comparison and
Wald-tests for testing individual (or groups of) parameters. The
log-likelihood function for a set of independent observations $y_i$,
$i=1,2,\ldots,n$ from the COM-Poisson$_\mu$ distribution has the
following form,
\begin{equation}
  \label{eqn:ll-rcmp}
  \ell = \ell(\bm{\beta}, \phi \mid \bm{y}) =
  e^\phi \left [
    \sum_{i=1}^n y_i
    \log \left( \mu_i + \frac{e^\phi-1}{2e^\phi} \right ) -
    \sum_{i=1}^n \log(y_i!) \right ] -
  \sum_{i=1}^n \log(Z(\mu_i, \phi)),
\end{equation}
where $\mu_i = \exp(\bm{x}_i^\top\bm{\beta})$, with
$\bm{x}_i^\top = (x_{i1},\, x_{i2},\, \ldots,\, x_{ip})$ is a vector of
known covariates for the $i$-th observation, and $(\bm{\beta},\, \phi)
\in \mathbb{R}^{p+1}$. The normalizing constant $Z(\mu_i, \phi)$ is
given by
\begin{equation}
\label{eqn:infseries}
  Z(\mu_i, \phi) = \sum_{j=0}^\infty \left [ \left (
    \mu_i + \frac{e^\phi - 1}{2e^\phi} \right )^{je^\phi}
  \frac{1}{(j!)^{e^\phi}} \right ].
\end{equation}
The evaluation of the log-likelihood function for each observation
involves the computation of the infinite series
(\autoref{eqn:infseries}).  Thus, the fitting procedure is computationally
expensive for regions of the parameter space where the convergence of
the infinite sum is slow.

Parameter estimation requires the numerical maximization of
\autoref{eqn:ll-rcmp}. Since the derivatives of $\ell$ cannot be
obtained in closed forms, we use the \texttt{BFGS}
algorithm~\citep{Nocedal1995} as implemented in the function
\texttt{optim()} in \texttt{R} \citep{Rcore2017}. Standard errors for
the regression coefficients are obtained based on the observed
information matrix $\mathcal{I}(\bm{\theta})$, where
$\mathcal{I}(\bm{\theta}) = -\mathcal{H}(\bm{\theta})$ (Hessian matrix)
is computed numerically by central finite differences. Standard errors
for $\hat{\mu}_i$ and hence confidence intervals are obtained by using
the delta method~\citep[p. 89]{Pawitan2001}.

Parameter estimation for the COM-Poisson regression model in the
original parametrization is analogous to that presented for the
COM-Poisson$_\mu$ distribution, however, it uses \autoref{eqn:ll-rcmp}
re-expressed in terms of $\lambda$. Here, even for the standard
COM-Poisson distribution, the dispersion parameter is taken on the log
scale to avoid numerical issues.

For the applications we also fitted the quasi-Poisson model
\citep{Wedderburn1974} as a baseline model. This approach is based only
on a second-moment assumption and without specific underlying probablity
model is less restrictive. In this model the variance of the response
variable is specified by an additional dispersion parameter $\sigma$,
with $\textrm{Var}(Y_i)=\sigma \mu_i$. These models are fitted in
\texttt{R} using the function \texttt{glm(..., family = quasipoisson)}.

\section{Simulation study}
\label{simulation-study}

In this section we report a simulation study to assess the properties of
the maximum likelihood estimators, the approximate parameter
orthogonality of the reparametrized model, as well as the flexibility of
the COM-Poisson regression model to deal with non-equidispersed count
data.

We considered average counts varying from $3$ to $27$ arising from a
regression model with a continuous and a categorical covariate. The
continuous covariate~($\bm{x}_1$) was generated as a linearly increasing
sequence from $0$ to $1$ with length equal to the sample size.
Similarly, the categorical covariate~($\bm{x}_2$) was generated as a
sequence of three values each one repeated $n/3$ times (rounding up when
required), where $n$ denotes the sample size. The parameter $\mu$ of the
reparametrized COM-Poisson random variable is given by
$\bm{\mu} = \exp(\beta_0 + \beta_1 \bm{x}_1 + \beta_{21} \bm{x}_{21} +
\beta_{22} \bm{x}_{22})$,
where $\bm{x}_{21}$ and $\bm{x}_{22}$ are dummy representing the levels
of $\bm{x}_2$.  The regression coefficients were fixed at the values,
$\beta_0 = 2$, $\beta_1 = 0.5$, $\beta_{21} = 0.8$ and
$\beta_{22} = -0.8$.

We designed four simulation scenarios by considering different values of
the dispersion parameter $\phi = -1.6, -1.0, 0.0$ and $1.8$.  Thus, we
have strong and moderate overdispersion, equidispersion, and
underdispersion, respectively.  \autoref{fig:justpars} shows the
variation of the average counts (left) and dispersion index (right) for
each value of the dispersion parameter considered in the simulation
study.  These configurations allow us to assess the properties of the
maximum likelihood estimators not only in extreme situations, such as
high counts and low dispersion, and low counts and high dispersion, but
also in the standard case of equidispersion.

<<load-simulation, cache=TRUE>>=
@

<<justpars, fig.height=4, fig.width=9, fig.cap="Average counts (left) and dispersion indexes (right) for each scenario considered in the simulation study.">>=
@

In order to check the consistency of the estimators we considered four
different sample sizes: $50$, $100$, $300$ and $1000$; generating $1000$
data sets in each case. For sample sizes 50 and 100, we have 29 and 8
simulations of the 1000 where the fitting algorithm did not
converge. These non-convergence situations occurred for $\phi=-1.6$. In
\autoref{fig:bias-plot}, we show the bias of the estimators for each
simulation scenario (combination between values of the dispersion
parameter and samples sizes) along with the confidence intervals
calculated as average bias plus and minus $1.96$ times the average
standard error. The scales are standardized for each parameter by
dividing the average bias by the average standard error obtained for the
sample of size $50$.

<<bias-data, cache=TRUE>>=
@

The results in \autoref{fig:bias-plot} show that for all dispersion
levels, both the average bias and standard errors tend to $0$ as the
sample size increases. Thus the estimators for the regression parameters
are unbiased, consistent and their empirical distributions are
symmetric. For the dispersion parameter, the estimator is asymptotically
unbiased; in small samples the parameter is overestimated and the
empirical distribution is slightly right-skewed.

<<bias-plot, cache=TRUE, fig.height=6.5, fig.width=8.5, fig.cap="Distributions of standardized bias (gray box-plots) and average with confidence intervals (black segments) by different sample sizes and dispersion levels.">>=
@

<<coverage-data>>=
@

\autoref{fig:coverage-plot} presents the empirical coverage rate of the
asymptotic confidence intervals. The results show that for the
regression parameters the empirical coverage rates are close to the
nominal level of 95\% for sample sizes greater than $100$ and all
simulation scenarios. For the dispersion parameter the empirical
coverage rates are slightly lower than the nominal level; however, they
become closer to the nominal level for large samples.  The worst
scenario is when we have a small sample size and strong overdispersed
counts.

<<coverage-plot, fig.height=4, fig.width=9, fig.cap="Coverage rate based on confidence intervals obtained by quadratic approximation for different sample sizes and dispersion levels.">>=
@

To check the orthogonality property we compute the covariance matrix
between maximum likelihood estimators
$\hat{\bm{\theta}} = (\hat{\bm{\beta}}, \phi)$, obtained from the
observed information matrix,
Cov$(\hat{\bm{\theta}}) = \mathcal{I}^{-1}(\bm{\theta})$.
\autoref{fig:ortho-plot} shows the correlation between the regression
and dispersion parameter estimators for each simulation scenario, on the
correlation scale. The correlations are gnerally close to zero in all
cases suggesting the orthogonality property for the reparametrized
model. Interestingly, results in the first panel show that
corr($\hat{\beta}_{22}, \hat{\phi}$) is not very close to zero (values
between $-0.4$ and $0.2$) for strong overdispersion ($\phi = -1.6$).

<<ortho-plot, fig.height=4.5, fig.width=9, fig.cap="Empirical correlations between regression and dispersion parameters by different sample sizes and dispersion levels.">>=
@

<<ortho-surf, fig.height=5, fig.width=9, fig.cap="Deviance surfaces contour plots under original and proposed parametrization for four simulated data sets of size 1000 with different dispersion parameters. The ellipses are confidence regions (90, 95 and 99\\%), dotted lines are the maximum likelihood estimates, and points are the real parameters used in the simulation.">>=
@

To illustrate the orthogonality, \autoref{fig:ortho-surf} displays
contour plots of the deviance surfaces for four simulated data sets of
size 1000 with $\mu=5$ and different values of the dispersion
parameters. The shapes of the deviance function show that the proposed
parametrization is better for both computation and asymptotic
(normal-based) inference. Furthermore, it is interesting to note that
the deviance function shape under strong overdispersion ($\phi=-1.6$) is
not as well behaved as the others; this is due to the inaccuracy of the
mean approximation \autoref{fig:approx-plot}(a). Although the
reparametrized model is always valid, $\mu$ and $\phi$ are orthogonal
only when $\mu$ is the expectation of the distribution, in other words,
when the approximation $\lambda - (\nu-1)/(2\nu)$ is accurate. This also
explains the results of corr$(\hat{\beta}_{22}, \phi)$ in the first
panel of \autoref{fig:ortho-plot}, since $\beta_{22}$ is negative and
associated with low counts.

\section{Case studies}
\label{case-studies}

In this section, we describe three illustrative examples of count data
analysis. For the analyses as alternative models we considered the
standard Poisson regression model, the COM-Poisson model in the two
forms (original and new parametrization), and the quasi-Poisson
regression model.  The data sets and \texttt{R} code for their analysis
are available as supplementary material.

\subsection{Artificial defoliation in cotton phenology}
\label{case-cotton}

This example relates to cotton plants (\textit{Gossypium hirsutum})
subjected to five levels of artificial defoliation (\texttt{des}) at
each of five different growth stages (\texttt{est}). The main goal of
this study was to assess the effect of defoliation levels at the
different growth stages on the cotton production, measured by the number
of bolls produced. The experimental study was conducted in a greenhouse
and used a completely randomized design with five replicates. This data
set was previously analyzed by~\citet{Zeviani2014} using the Gamma-Count
distribution.

Following~\citet{Zeviani2014}, the linear predictor is given by
$$\eta_i = \beta_0 + \beta_{1j} \texttt{def}_i + \beta_{2j}
\texttt{def}_i^2$$
where $\eta_i = \log(\lambda_i)$ for the original COM-Poisson and
$\eta_i = \log(\mu_{ij})$ for COM-Poisson$_\mu$. The parameter
$\mu_{ij}$ is the (approximated) expected number of cotton bolls for the
$i$-th defoliation level ($i=$ 1: 0\%, 2: 25\%, 3: 50\%, 4: 75\% e 5:
100\%) and $j$-th growth stage ($j$ = 1: vegetative, 2: flower bud, 3:
blossom, 4: boll, 5: boll open), i.e. we have a different second order
effect of defoliation at each growth stage. The parameter estimates and
goodness-of-fit measures for the Poisson, COM-Poisson,
COM-Poisson$_\mu$, and quasi-Poisson regression models are presented in
\autoref{tab:coef-cotton}.

<<fit-cotton, cache=TRUE, include=FALSE>>=
@

\begin{table}[!ht]
\centering \small
\caption{Parameter estimates (Est) and ratio between estimate and
  standard error (SE) for the four model strategies for the analysis
  of the cotton experiment.}
\label{tab:coef-cotton}
\begin{tabular}{lrrrrrrrr}
  \toprule
  & \multicolumn{2}{c}{Poisson} &
    \multicolumn{2}{c}{COM-Poisson} &
    \multicolumn{2}{c}{COM-Poisson$_\mu$} &
    \multicolumn{2}{c}{Quasi-Poisson} \\
\cmidrule(lr){2-3} \cmidrule(lr){4-5} \cmidrule(lr){6-7} \cmidrule(lr){8-9}
<<results-cotton, results="asis">>=
@
\end{tabular}
\end{table}

The results presented in \autoref{tab:coef-cotton} show that the
goodness-of-fit measures (log-likelihood, AIC and BIC) are quite similar
for the COM-Poisson and COM-Poisson$_\mu$ models. This suggests that the
reparametrization does not change the model fit, as is to be expected.
The Poisson model is clearly unsuitable, being overly conservative, due
to the overestimated standard errors.  The difference in terms of
$-2\times$log-likelihood value from the Poisson to the COM-Poisson$_\mu$
model is $94.811$ for one additional parameter, which confirms the
significantly improved fit of the COM-Poisson$_{\mu}$ model.  Finally,
the estimated value of the dispersion parameter $\hat{\phi} = 1.582$
suggests substantial underdispersion.

Furthermore, results in \autoref{tab:coef-cotton} also show the
advantage of the COM-Poisson$_\mu$ model, since the regression parameter
estimates are quite similar to those obtained for the Poisson model,
whereas estimates from the original COM-Poisson model are on a different
and not easily interpreted scale.  The ratios between estimates and
their respective standard errors for the two COM-Poisson models are very
close to ratios from the y quasi-Poisson model. However, it is important
to note that COM-Poisson model is a full parametric approach, i.e. there
is a probability distribution associated to the counts, while the
quasi-Poisson model is based only on second-moment assumptions.

\autoref{fig:pred-cotton} presents the observed and fitted values with
confidence intervals (95\%) as a function of the defoliation level for
each growth stage. The fitted values are the same for the Poisson and
COM-Poisson$_{\mu}$ models, however, the confidence intervals are larger
for the Poisson model because of the inappropriate equidispersion
assumption.  The results from the COM-Poisson$_{\mu}$ model are
consistent with those from the Gamma-Count model~\citep{Zeviani2014},
Poisson-Tweedie~\citep{Bonat2017} and the alternative parametrization of
the COM-Poisson distribution proposed by~\citet{Huang2017}.  All of
these models indicated underdispersion and significant effects of
defoliation at the vegetative, blossom and boll growth stages.

<<pred-cotton, fig.height=3.5, fig.width=8.5, fig.cap="Scatterplots of the observed data and curves of fitted values with 95\\% confidence intervals as functions of the defoliation level for each growth stage.">>=
@

In order to assess the relation between $\bm{\mu}$ and $\phi$ in the
COM-Poisson$_\mu$ parametrization, \autoref{tab:corr-cotton} presents
the empirical correlations between the regression and dispersion
parameters, as computed by the asymptotic covariance matrix of the
estimates, i.e. the inverse of the observed information.  The
correlations are practically zero for the COM-Poisson$_\mu$, however,
with the original parametrization some correlations are quite large, in
particular for the parameter $\beta_0$ (due to the effects
parametrization in the linear predictor).  This result explains the
improved performance of the maximization algorithm in the new
parametrization.  It is also important to note that as initial
regression parameter values for the \texttt{BFGS} algorithm are provided
by the Poisson model, then in the COM-Poisson$_{\mu}$ model the initial
values are very close to the maximum likelihood estimates and
maximization effort is essentially only on the dispersion parameter
$\phi$. To compare the computational times for the two parametrizations
we repeat the fitting $50$ times. In this example the COM-Poisson$_\mu$
fit was, on average, $38$\% faster than for the original version.

<<corr-cotton, results="asis">>=
@

\begin{table}[ht]
\centering
\caption{Empirical correlations between $\hat{\phi}$ and
  $\hat{\bm{\beta}}$ for the two parametrizations of COM-Poisson model
  fit to underdispersed data.}
\label{tab:corr-cotton}
\begingroup\small
\begin{tabular}{lrrrrrr}
  \toprule
<<cres-cotton, results="asis">>=
@
\end{tabular}
\endgroup
\end{table}

\subsection{Soil moisture and potassium doses on soybean culture}
\label{case-soybean}

In this second example we consider a study of potassium doses and soil
moisture levels on soybean (\emph{Glicine Max}) production. This was set
up as a $5\times 3$ factorial experiment in a randomized complete block
design with 5 replicates.  Five different potassium doses (\texttt{K})
(0, 0.3, 0.6, 1.2 and 1.8 $\times$ 100mg dm$^{-3}$) were applied to the
soil and soil moisture (\texttt{umid}) levels were controlled at (37.5,
50, 62.5\%) . The experiment was carried out in a greenhouse and the
experimental units were pots with two plants in each. The count response
measured was the total number of bean seeds per pot~\citep{Serafim2012}.
\autoref{fig:desc-soy} (left) shows the number of bean seeds recorded
for each combination of potassium dose and moisture level; it is
important to note the indication of a quadratic effect of the potassium
levels, as indicated by the smoothers. Most points in the sample
variance \textit{versus} sample means dispersion diagram (right) are
above the identity line, suggesting overdispersion (block effect not yet
removed).

<<fit-soy, cache=TRUE, include=FALSE>>=
@

<<desc-soy, fig.height=4, fig.width=8, fig.cap="Number of bean seeds per pot for each potassium dose and moisture level (left) and sample mean against sample variance of the five replicates for each experimental treatment (right). Solid lines are smoothing curves (left) and the least-squares line (right).">>=
@

For the analysis of this data set, based on the descriptive analysis
(\autoref{fig:desc-soy}), we proposed the following linear predictor
$$
\eta_i = \beta_0 + \gamma_i + \tau_j +
  \beta_{1}\texttt{K}_k + \beta_{2}\texttt{K}_k^2 +
  \beta_{3j}\texttt{K}_k
$$
where $\gamma_i$ is the effect of $i$-th block ($i=$1: block II, 2:
block III, 3: block IV and 4: block V), $\tau_j$ is the effect of $j$-th
moisture level ($j=$1: 50\% and 2: 62.5\%), $\beta_{1}$ and $\beta_{2}$
give the baseline quadratic response over potassium levels \texttt{K}
($k=$1: 0.0, 2: 0.3, 3: 0.6, 4: 1.2, 5: 1.8 100mg dm$^{-3}$) and
$\beta_{3j}$ is interaction of the first order potassium effect
(\texttt{K}) for the $j$-th moisture level (\texttt{umid}); The
predictor $\eta_i$ is $\log(\lambda_i)$ for the original COM-Poisson and
$\log(\mu_{ij})$ for COM-Poisson$_\mu$. \autoref{tab:coef-soy} presents
the estimates, ratio between estimate and standard error, and
goodness-of-fit measures for the alternative models.

The results in \autoref{tab:coef-soy} show that the two parametrizations
of the COM-Poisson model presented very similar goodness-of-fit measures
and a better fit than the Poisson model. The $-2\times$difference
between the log-likelihoods of the Poisson and COM-Poisson models was
$29.697$ on 1 degree of freedom, indicating that $\phi$ is significantly
different from zero. The estimate of $\phi$ ($-0.782$) indicates
overdispersion, corroborating the descriptive analysis. Concerning the
regression parameters, the similarities between the models are analogous
to those in the previous section. Both models show effects of block,
potassium dose and moisture level, however the Poisson model indicates
the effects with greater significance, because it does not take account
of the extra variability and so underestimates standard errors.

\begin{table}[!ht]
\centering \small
\caption{Parameter estimates (Est) and ratio between estimate and
  standard error (SE) for the four model strategies for the analysis of
  the soybean experiment.}
\label{tab:coef-soy}
\begin{tabular}{lrrrrrrrr}
  \toprule
  & \multicolumn{2}{c}{Poisson} &
    \multicolumn{2}{c}{COM-Poisson} &
    \multicolumn{2}{c}{COM-Poisson$_\mu$} &
    \multicolumn{2}{c}{Quasi-Poisson} \\
\cmidrule(lr){2-3} \cmidrule(lr){4-5} \cmidrule(lr){6-7} \cmidrule(lr){8-9}
<<results-soya, results="asis">>=
@
\end{tabular}
\end{table}

The infinite sum $Z(\mu, \phi)$ in the cases of overdispersed count data
requires a larger upper bound to reach convergence. Thus, in these cases
the computation of the log-likelihood function is computationally
expensive.  For the data set considered here , the upper bound was fixed
at $700$.  To reach convergence the \texttt{BFGS} algorithm evaluated
the log-likelihood function 264 times when using the original
parametrization of the COM-Poisson distribution and only 20 times with
the new parametrization.  In terms of computational time, for 50
repetitions of the fit, the proposed reparametrization was, on average,
$110\%$ faster than the original one.  This is probably due to the
better behaviour of the log-likelihood function as well as more relevant
initial values from the Poisson fit. In \autoref{tab:corr-soy}, we
present the empirical correlations between the regression and dispersion
parameter estimates. Again, the correlations are close to zero for the
COM-Poisson$_\mu$ model, indicating the empirical orthogonality between
$\mu$ and $\phi$.

<<corr-soy, results="asis">>=
@

\begin{table}[ht]
\centering
\caption{Empirical correlations between $\hat{\phi}$ and
  $\hat{\bm{\beta}}$ for the two parametrizations of COM-Poisson model
  fit to overdispersed data.}
\label{tab:corr-soy}
\begingroup\small
\begin{tabular}{lrrrrrrr}
  \toprule
<<cres-soy, results="asis">>=
@
\end{tabular}
\endgroup
\end{table}

The observed and fitted counts for each humidity level with confidence
intervals are shown in \autoref{fig:pred-soy}. The fitted values are
identical for the Poisson and COM-Poisson$_{\mu}$ models, leading to the
same conclusions. On the other hand, confidence intervals for the
Poisson model are narrower than those from the COM-Poisson$_{\mu}$, due
to the inappropriate equidispersion assumption of the Poisson model. The
confidence intervals from the COM-Poisson$_{\mu}$ and quasi-Poisson
models are also very similar, which again shows the already highlighted
similarity between these approaches, however only the COM-Poisson
model$_{\mu}$ corresponds to a fully specified probability model.

<<pred-soy, fig.height=3.8, fig.width=8, fig.cap="Observed bean seeds counts as functions of potassium doses and soil humidity levels with fitted curves and confidence intervals (95\\%).">>=
@

\subsection{Assessing toxicity of nitrofen in aquatic systems}

Nitrofen is a herbicide that was used extensively for the control of
broad-leaved and grass weeds in cereals and rice. Although it is
relatively non-toxic to adult mammals, nitrofen is a significant
tetragen and mutagen. It is also acutely toxic and reproductively toxic
to cladoceran zooplankton. Nitrofen is no longer in commercial use in
the U.S., having been the first pesticide to be withdrawn due to
tetragenic effects~\citep{Bailer1994}.

This data set comes from an experiment to measure the reproductive
toxicity of the herbicide, nitrofen, on a species of zooplankton
(\textit{Ceriodaphnia dubia}). Fifty animals were randomized into
batches of ten and each batch was put in a solution with a measured
concentration of nitrofen ($0, 0.8, 1.6, 2.35$ and $3.10$
$\mu$g$/10^2$litre) (\texttt{dose}). Subsequently, the number of live
offspring was recorded.

<<fit-ovos, cahe=TRUE, include=FALSE>>=
@

For this data set we consider three models with linear predictors,
\begin{center}
\begin{minipage}{12cm}
Linear: $\eta_i = \beta_0 + \beta_1 \texttt{dose}_i$\\
Quadratic: $\eta_i = \beta_0 + \beta_1 \texttt{dose}_i +
             \beta_2 \texttt{dose}_i^2$\\
Cubic: $\eta_i = \beta_0 + \beta_1 \texttt{dose}_i +
             \beta_2 \texttt{dose}_i^2 + \beta_3 \texttt{dose}_i^3$.
\end{minipage}
\end{center}
where $\eta_i = \log(\lambda_i)$ for the original COM-Poisson and
$\eta_i = \log(\mu_{ij})$ for COM-Poisson$_\mu$.

\begin{table}[!ht]
\centering \small
\caption{Model fit measures and comparisons between predictors and
  models fitted to the nitrofen data.}
\label{tab:anova-ovos}
\begin{tabularx}{\textwidth}{lCCCCCrC}
  \toprule
 Poisson & np & $\ell$ & AIC & 2(diff $\ell$) & diff np & P($>\rchi^2$) & \\
 \midrule
<<anova-ovos1, results="asis">>=
@
\specialrule{0em}{0.5em}{0em} %% Apenas para espaçamento
  COM-Poisson & np & $\ell$ & AIC & 2(diff $\ell$) & diff np &
  P($>\rchi^2$) & $\hat{\phi}$ \\
  \midrule
<<anova-ovos2, results="asis">>=
@
\specialrule{0em}{0.5em}{0em} %% Apenas para espaçamento
  COM-Poisson$_\mu$ & np & $\ell$ & AIC & 2(diff $\ell$) & diff np &
  P($>\rchi^2$) & $\hat{\phi}$ \\
  \midrule
<<anova-ovos3, results="asis">>=
@
\specialrule{0em}{0.5em}{0em} %% Apenas para espaçamento
  Quasi-Poisson & np & QDev & AIC & F & diff np & P($>F$) & $\hat{\sigma}$ \\
  \midrule
<<anova-ovos4, results="asis">>=
@
 \bottomrule
\end{tabularx}
%\vspace{-1mm}

\footnotesize \raggedright $\#p$, number of parameters; diff $\ell$,
difference in log-likelihoods; QDev, quasi-deviance, F, F statistics
based on quasi-deviances; diff $\#p$, difference in number of parameters.
\end{table}

\autoref{tab:anova-ovos} summarizes the results of the fitted models and
likelihood ratio tests comparing the sequence of predictors. All models
indicate the significance of the cubic effect of the nitrofen
concentration.  Considering this predictor, there is evidence of
equidispersed counts, the $\phi$ estimate of the COM-Poisson is close to
zero and the $\sigma$ estimate of the quasi-Poisson is close to one.  It
is interesting to note that if we omit the higher order effects the
models show evidence of overdispersion --- this exemplifies the
discussion on the possible causes of overdispersion in
\autoref{introduction}.  We also note that the quasi-Poisson approach,
although robust to the equidispersion assumption, shows higher
descriptive levels ($p$-values) than parametric models, that is, the
tests under parametric models are more powerful than the ones under the
quasi-Poisson model in the equidispersed case.

In \autoref{tab:coef-ovos}, we present the estimates of the regression
parameters for the cubic dose models. The interpretations are similar to
the other case studies, however, here the Poisson model is also suitable
for indicating the significance of the covariate effects.  In addition,
note that the parameter estimates of the original COM-Poisson model are
comparable to the others models. This occurs because we are in the
particular case, where $\phi \approx 0$, which implies
$\lambda \approx \mu$.

\begin{table}[!ht]
\centering
\caption{Parameter estimates (Est) and ratio between estimate and
  standard error (SE) for the four model strategies for the analysis of
  the nitrofen experiment.}
\label{tab:coef-ovos}
\begin{tabular}{lrrrrrrrr}
  \toprule
  & \multicolumn{2}{c}{Poisson} &
    \multicolumn{2}{c}{COM-Poisson} &
    \multicolumn{2}{c}{COM-Poisson$_\mu$} &
    \multicolumn{2}{c}{Quasi-Poisson} \\
\cmidrule(lr){2-3} \cmidrule(lr){4-5} \cmidrule(lr){6-7} \cmidrule(lr){8-9}
<<coef-ovos, results="asis">>=
@
\end{tabular}
\end{table}

\autoref{fig:pred-ovos} shows the number of live off-spring observed and
fitted curves along with confidence intervals for the different cubic
dose models.  The fitted values and confidence intervals are identical
and have a complete overlap. This shows that the estimation of the extra
dispersion parameter does not affect the estimation of the regression
coefficients in the case of equidispersed counts.

<<pred-ovos, fig.height=3.5, fig.width=5.5, out.width="0.7\\textwidth", fig.cap="Number of live offspring observed for each nitrofen concentration level with fitted curves and 95\\% confidence intervals.">>=
@

Finally, in \autoref{tab:corr-ovos} we present the empirical
correlations between the regression and dispersion parameter estimates.
The results show that even in the special case ($\phi=0$), the empirical
correlations for the original COM-Poisson model are not zero.  For the
reparametrized model, as discussed in the previous sections, the
correlations are practically null. The computational times for fifty
repetitions of fit are similar; the average time to fit the
COM-Poisson$_\mu$ and COM-Poisson models is 1.19 and 1.09 seconds,
respectively.

<<corr-ovos, results="asis">>=
@

\section{Concluding remarks}
\label{conclusion}

In this paper, we presented and characterized a novel reparametrization
of the COM-Poisson distribution and associated regression model. The
reparametrization was based on a simple asymptotic approximation for the
expectation of the COM-Poisson distribution. The main advantage of the
proposed reparametrization is the simple interpretation of the
regression coefficients in terms of the expectation of the response
variable as usual in the generalized linear models context. Thus, it is
possible to compare directly the results from the COM-Poisson model with
those from standard approaches, such as the Poisson and quasi-Poisson
regression models. Furthermore, in the new parametrization the
COM-Poisson distribution is indexed by $\mu$ and an extra dispersion
parameter $\phi$ which our studies suggest are approximately
orthogonal. Overall the approach is similar to \Citeauthor{Huang2017}'s
(\citeyear{Huang2017}) parametrization but ours is simpler, because the
$\mu$ used here is obtained from simple algebra.

The proposed parametrization is always valid, only the interpretation of
$\mu$ parameter, as the expectation of the distribution, depends on the
accuracy of the approximation. We evaluated the accuracy of the
approximations for the expectation and variance of the COM-Poisson
distribution by considering quadratic approximation errors.  The results
showed that the approximation for the expectation is accurate for a
large part of the parameter space, supporting our reparametrization. We
discuss the properties and flexibility of the distribution to deal with
count data through considerations of dispersion, zero-inflation and
heavy-tail indexes. A simulation study was used to assess the properties
of the reparametrized COM-Poisson model and its ability to deal with
different levels of dispersion, as well as the properties of the maximum
likelihood estimators. The results of our simulation study suggested
that the maximum likelihood estimators of the regression and dispersion
parameters are unbiased and consistent.  The empirical coverage rates of
the confidence intervals computed based on the asymptotic distribution
of the maximum likelihood estimators are close to the nominal level for
sample sizes greater than $100$.  The worst case scenario is when we
have a small sample size and strongly overdispersed counts.  In general,
we recommend the use of the asymptotic confidence intervals for
computational simplicity, although of course bootstrap intervals could
be obtained, but would involve extensive computation.

The three examples and data analyses have shown that the COM-Poisson
regression model is a suitable choice to deal with generally dispersed
count data, equi-, under- and over-dispersed.  The observed empirical
correlation between the regression and dispersion parameter estimators
and deviance surfaces suggest approximate orthogonality between $\mu$
and $\phi$ in the COM-Poisson$_\mu$ distribution. Thus, the
computational procedure based on the proposed reparametrization is
faster than that for the original parametrization.

In general, the results presented by the reparametrized COM-Poisson
models were satisfactory and comparable to the conventional approaches.
Therefore, its use in the analysis of count data is encouraged. The
computational routines for fitting the original and reparametrized
COM-Poisson regression models are available in the supplementary
material\textsuperscript{\ref{papercompanion}}.

There are many possible extensions to the model discussed in this paper,
including simulation studies to assess the model robustness against
model misspecification and to assess the theoretical approximations for
$Z(\lambda, \nu)$ (or $Z(\mu,\phi))$ \citep{Gaunt2017}. Another simple
extension of the proposed model is to model both $\mu$ and $\phi$
parameters as functions of covariates in a double generalized linear
model framework.  Finally, the reparametrized version of the COM-Poisson
model would also be highly suitable for the specification of generalized
linear mixed COM-Poisson models.

%-----------------------------------------------------------------------
\bibliography{references.bib}

\end{document}